\label{ch11:acknowledgements}

\section{Personal acknowledgements}\label{personal-acknowledgements}

This thesis would not have been possible without the tremendous support I have been so incredibly lucky to get from my closest and dearest throughout my PhD journey.

First of all I must thank my wonderful and amazing wife Gaby. You are my strongest pillar, making me motivated and focused, supporting me even when I have been difficult or absent-minded. I could not have done this without your love, and I am forever grateful. ¡Te amo con todo mi corazón! 

Much of my daily joy comes from the my children,
and I am glad for all the time we have together -- including during COVID-19 lockdown when work, study and life mixed all up! (apologies to readers if you find any Odd Squad references). You will probably insist I also thank Waldo -- OK, he has also been there, although mostly sleeping in his little corner!

I want to thank my family in Norway -- specially my parents who have been encouraging me throughout, and my siblings and cousins who always make it a pleasure to ``come home'' and give me regular updates.  Thank you to all my friends, in particular Laura, Juan, Ian, Sverre, Erlend, and Magnus. 

I am forever in debt to Paul Groth and I want to thank you for agreeing to supervise me as your PhD student, which must have been challenging at the best of times. Thank you for all your patience, persistence and for your pragmatic directions that kept me on track. Thanks also to the \footurl{https://indelab.org/people/}{INDE lab} at University of Amsterdam, particularly for the daily colleaguality we got during lockdown -- I wish I could have spent more time with you in person!

Thank you to Carole Goble, you have always been supporting me in so many ways since I started working for you in 2006, and you convinced me to start this PhD in the first place! No wonder you kept asking if I was finished yet! I appreciate our friendship and how you have mentored me to grow from a geeky developer to an acedemic. I hope we continue our exciting collaborations as ``proper'' colleagues!

I am so happy to have worked at the \footurl{https://esciencelab.org.uk/people/}{eScience Lab} at The University of Manchester for all these years and during this PhD. I particularly want to thank Shoaib Sufi for your wisdom and our deep conversations, not to mention your reliability whenever I have been stunned by bureaucracy or other struggles -- I've probably been the oddest person for you to line manage! 
Thank you to all my amazing colleagues at the eScience Lab -- specially Stuart, Finn, Alex, Nick, Munazah, Alan -- you are a pleasure to work with, not just for your skills and expertise, but also your friendship and support.

I want to thank Meznah, you have been in the odd position of being co-supervised by me for your own PhD, and at the same time pushing me to finish mine! You bring such a joy with everything you do!  Thanks to my fellow PhD and EngD students in Manchester: Fuqi, Ebitsam, Yo  -- you inspire me! Thanks to Rudolf Wittner, your productive provenance visits to Manchester have hopefully helped your PhD as much as mine!  

I also want to thank Andrew Stewart, Uli Sattler and the rest of Department of Computer Science, who have been so accommodating, and you are now welcoming me again as I step into the academic world fully. It's not like you didn't prepare me!
I also want to thank the UCU union members in Manchester \footurl{https://manchester.web.ucu.org.uk/}{(UMUCU)} -- particularly Bijan and Ben for their personal support. I may have lost track of how many times we've been on the picket line during this PhD (no, I didn't manage to do much writing on strike days), but you have also showed me not to give up!  Thank you to the engaging University of Manchester Students' Union, for standing up for what is right!


\section{Community acknowledgements}\label{community-acknowledgements}

I am grateful for all the wonderful discussions, technical contributrions and long-lasting friendships formed in the RO-Crate community (listed \vpageref{ro-crate-community}) since its early inception \cite{Ó Carragáin 2019a} at Workshop on Research Objects \footurl{https://www.researchobject.org/ro2018/}{(RO2018)}. Without you there would not have been any RO-Crate!

I would particularly like to thank Peter Sefton, without which co-chairing, enthusiasm and pragmatism I would easily had got lost the weeds.  I am forever in debt to Eoghan Ó Carragáin who managed to get me and Peter to agree even on the trickiest issues, and led the community through its formative years.

I want to thank Simone Leo, which persistent work and leadership on \footurl{https://w3id.org/ro/wfrun/}{Workflow Run Crate} and the RO-Crate Python library \cite{De Geest 2023a} has shown that RO-Crate works in practice even for detailed provenance. 

Thank you to Bruno Kinoshita who kindly (and on own initiative) helped proof this whole thesis, you have contributed to our common goals in so many ways, going back to when we first worked on Semantic Web, Apache Jena, and Common Workflow Language and later RO-Crate.

I am very grateful to Daniel Garijo, with whom I have enjoyed many insightful discussions and idea developments for more than a decade (since the early days of Research Objects!), from the cellar of Dagstuhl to beaches of San Diego! I appreciate your enthusiasm and many invaluable and constructive contributions to manuscripts, specifications and code.

I want to thank Leyla Jael Castro, for your continuous positivity and drive, and for pushing us to run RO-Crate training workshops together with Núria Queralt Rosinach and Claus Weiland. I always appreciate our discussions and writing together, and I am so glad that you are much better than me to remember submission deadlines!

RO-Crate is standing on the shoulders of giants, and I would like to thank the whole Research Object community \cite{Goble 2018} for persisting on the early RO ideas \cite{Newman 2009,Bechhofer 2013}. I have particular fond memories of whiteboard sessions with Sean Bechhofer, Khalid Belhajjame, Paolo Missier, David De Roure, and Kevin Page in the productive \footurl{https://s11.no/2020/archive/wf4ever/}{Wf4Ever project} that laid the foundation for what is now RO-Crate.

This PhD is in many ways born out of 
\footurl{https://elixir-europe.org/}{ELIXIR Europe}, not just 
through its many fruitful co-author contributions 
(section \vref{ch10:my-contributions}), 
but also as the ``ELIXIR way'' of working is emblematic for the approach taken by RO-Crate community: open collaborations where instutional and project boundaries are blurred out in favour of jointly building pragmatic solutions.

I would like to thank the lovely people of ELIXIR Belgium (incl. Frederik, Ignacio, Paul, Rafael, Bert), ELIXIR Spain (incl. José Mª, Laura, Salva, Adam, Pau), ELIXIR HUB (incl. Justin, Jonathan, Marieke, Gavin, Niklas, Katharina), ELIXIR DE (incl. Björn, Anika, Sebastian, Nils, Michael), ELIXIR CH (Alex, Patricia), ELIXIR UK (incl. Susanna, Nicola, Xenia, Tim, Neil), ELIXIR NL (Núria, Egon, Chris, Marco, Rob, Helena) and many others I may have missed listing. I want to particularly thank everyone who organized and participated in the \footurl{https://biohackathon-europe.org/}{ELIXIR Biohackathons} which always inspires me and has helped build our communities.

This thesis frequently mentions the Common Workflow Language\footnote{About 97 times!}, and many of the ideas spur out of methods that were first tried by the \footurl{https://www.commonwl.org/contributors/}{CWL community} (see for example \cite{Möller 2017,Robinson 2017,Khan 2019,Crusoe 2022}), which I am grateful for. I want to thank Michael Crusoe, not just for your tiredless work on bringing CWL together and willingness to put FAIR workflow theories into practice, but also for our enduring friendship and your support throughout most of the rabbit holes I've ventured into. I want to thank Peter Amstutz, 
Hervé Ménager and the rest of the CWL leadership team and the many volunteers.

I also need to thank members of the \footurl{https://biocomputeobject.org/}{BioCompute Object} community, particularly Hadley, Jonathon, Raja, Dennis, Vahan, Janisha, Amanda, and Nicola. 
I thank the \footurl{https://workflows.community/}{Workflows Community Initiative} in particular Sean R. Wilkinson, Rafael Ferreira da Silva, and Kyle Chard. 

From the \footurl{https://fairdo.org/}{FAIR Digital Object Forum} I need to place particular thanks to Peter Wittenburg, Luiz Olavo Bonino da Silva Santos, Maggie Hellström, Ulrich Schwardmann and Rainer Stotzka. I also want to thank all the members of the the \footurl{https://www.rd-alliance.org/}{Research Data Alliance}'s FAIR Digital Object Fabric interest group.

Mark Wilkinson, we first worked together on BioMoby services and Taverna workflows, which seems a long way from where we are today with FAIR! Looking back from this thesis, perhaps we could consider those to be early FAIR digital objects as well? I want to thank you for your enthusiasm,  and for bringing me in to the Apples2Apples hackathons -- thank you also to Robert, Richard, Wilko, David, Alban, and Allyson.

I am indebted to Herbert van de Sompel -- our conversations on the topic of aggregations, annotations, mementos and the Web started more than a decade ago; your clear ideas and powerful visions are at the heart of what this thesis tries to achieve. I am always energised by your talks, and truly appreciate working with you to promote FAIR Signposting and hope we will continue such exciting work together.


\subsection{RO-Crate Community}\label{communitylist}
\label{ro-crate-community}

As of 2024-01-08, the \emph{RO-Crate} \footurl{https://www.researchobject.org/ro-crate/community.html}{Community members} are:

\begin{longtable}{lll}
  Peter Sefton & \url{https://orcid.org/0000-0002-3545-944X} & (co-chair) 
\\
  Stian Soiland-Reyes & \url{https://orcid.org/0000-0001-9842-9718} & (co-chair) 
\\
  Eoghan Ó Carragáin & \url{https://orcid.org/0000-0001-8131-2150} & (emeritus chair) 
\\
  Oscar Corcho & \url{https://orcid.org/0000-0002-9260-0753}
\\
  Daniel Garijo & \url{https://orcid.org/0000-0003-0454-7145}
\\
  Raul Palma & \url{https://orcid.org/0000-0003-4289-4922}
\\
  Frederik Coppens & \url{https://orcid.org/0000-0001-6565-5145}
\\
  Carole Goble & \url{https://orcid.org/0000-0003-1219-2137}
\\
  José María Fernández & \url{https://orcid.org/0000-0002-4806-5140}
\\
  Kyle Chard & \url{https://orcid.org/0000-0002-7370-4805}
\\
  Jose Manuel Gomez-Perez & \url{https://orcid.org/0000-0002-5491-6431}
\\
  Michael R Crusoe & \url{https://orcid.org/0000-0002-2961-9670}
\\
  Ignacio Eguinoa & \url{https://orcid.org/0000-0002-6190-122X}
\\
  Nick Juty & \url{https://orcid.org/0000-0002-2036-8350}
\\
  Kristi Holmes & \url{https://orcid.org/0000-0001-8420-5254}
\\
  Jason A. Clark & \url{https://orcid.org/0000-0002-3588-6257}
\\
  Salvador Capella-Gutierrez & \url{https://orcid.org/0000-0002-0309-604X}
\\
  Alasdair J. G. Gray & \url{https://orcid.org/0000-0002-5711-4872}
\\
  Stuart Owen & \url{https://orcid.org/0000-0003-2130-0865}
\\
  Alan R Williams & \url{https://orcid.org/0000-0003-3156-2105}
\\
  Giacomo Tartari & \url{https://orcid.org/0000-0003-1130-2154}
\\
  Finn Bacall & \url{https://orcid.org/0000-0002-0048-3300}
\\
  Thomas Thelen & \url{https://orcid.org/0000-0002-1756-2128}
\\
  Hervé Ménager & \url{https://orcid.org/0000-0002-7552-1009}
\\
  Laura Rodríguez-Navas & \url{https://orcid.org/0000-0003-4929-1219}
\\
  Paul Walk & \url{https://orcid.org/0000-0003-1541-5631}
\\
  brandon whitehead & \url{https://orcid.org/0000-0002-0337-8610}
\\
  Mark Wilkinson & \url{https://orcid.org/0000-0001-6960-357X}
\\
  Paul Groth & \url{https://orcid.org/0000-0003-0183-6910}
\\
  Erich Bremer & \url{https://orcid.org/0000-0003-0223-1059}
\\
  LJ Garcia Castro & \url{https://orcid.org/0000-0003-3986-0510}
\\
  Karl Sebby & \url{https://orcid.org/0000-0001-6022-9825}
\\
  Alexander Kanitz & \url{https://orcid.org/0000-0002-3468-0652}
\\
  Ana Trisovic & \url{https://orcid.org/0000-0003-1991-0533}
\\
  Gavin Kennedy & \url{https://orcid.org/0000-0003-3910-0474}
\\
  Mark Graves & \url{https://orcid.org/0000-0003-3486-8193}
\\
  Jasper Koehorst & \url{https://orcid.org/0000-0001-8172-8981}
\\
  Simone Leo & \url{https://orcid.org/0000-0001-8271-5429}
\\
  Marc Portier & \url{https://orcid.org/0000-0002-9648-6484}
\\
  Paul Brack & \url{https://orcid.org/0000-0002-5432-2748}
\\
  Milan Ojsteršek & \url{https://orcid.org/0000-0003-1743-8300}
\\
  Bert Droesbeke & \url{https://orcid.org/0000-0003-0522-5674}
\\
  Chenxu Niu & \url{https://orcid.org/0000-0002-2142-1731}
\\
  Kosuke Tanabe & \url{https://orcid.org/0000-0002-9986-7223}
\\
  Tomasz Miksa & \url{https://orcid.org/0000-0002-4929-7875}
\\
  Marco La Rosa & \url{https://orcid.org/0000-0001-5383-6993}
\\
  Cedric Decruw & \url{https://orcid.org/0000-0001-6387-5988}
\\
  Andreas Czerniak & \url{https://orcid.org/0000-0003-3883-4169}
\\
  Jeremy Jay & \url{https://orcid.org/0000-0002-5761-7533}
\\
  Sergio Serra & \url{https://orcid.org/0000-0002-0792-8157}
\\
  Ronald Siebes & \url{https://orcid.org/0000-0001-8772-7904}
\\
  Shaun de Witt & \url{https://orcid.org/0000-0003-4196-3658}
\\
  Shady El Damaty & \url{https://orcid.org/0000-0002-2318-4477}
\\
  Douglas Lowe & \url{https://orcid.org/0000-0002-1248-3594}
\\
  Xuanqi Li & \url{https://orcid.org/0000-0003-1498-6205}
\\
  Sveinung Gundersen & \url{https://orcid.org/0000-0001-9888-7954}
\\
  Muhammad Radifar & \url{https://orcid.org/0000-0001-9156-9478}
\\
  Rudolf Wittner & \url{https://orcid.org/0000-0002-0003-2024}
\\
  Oliver Woolland & \url{https://orcid.org/0000-0002-4565-9760}
\\
  Paul De Geest & \url{https://orcid.org/0000-0002-8940-4946}
\\
  Douglas Fils & \url{https://orcid.org/0000-0002-2257-9127}
\\
  Florian Wetzels & \url{https://orcid.org/0000-0002-5526-7138}
\\
  Raül Sirvent & \url{https://orcid.org/0000-0003-0606-2512}
\\
  Abigail Miller & \url{https://orcid.org/0000-0001-9228-2882}
\\
  Jake Emerson & \url{https://orcid.org/0000-0003-0617-9219}
\\
  Davide Fucci & \url{https://orcid.org/0000-0002-0679-4361}
\\
  Bruno P. Kinoshita & \url{https://orcid.org/0000-0001-8250-4074}
\\
  Maciek Bąk & \url{https://orcid.org/0000-0003-1361-7301}  
\\
  Jens Hollunder & \url{https://orcid.org/0000-0003-3234-6762}
\\
  Martin Weise & \url{http://orcid.org/0000-0003-4216-302X}
\\
  Vartika Bisht & \url{https://orcid.org/0000-0002-1880-0597}
\\
  Toshiyuki Nishiyama Hiraki & \url{https://orcid.org/0000-0001-6712-6335}
\end{longtable}


\subsection{In remembrance}

I am saddened by the loss of \footurl{https://www.youtube.com/watch?v=y8EhezA5_LY}{Olav S. Bratli}, \footurl{Stu Allan}{https://stuallan.co.uk/}, \footurl{https://galaxyproject.org/jxtx/}{James Taylor}, \footurl{https://www.rd-alliance.org/remembering-sarah-jones}{Sarah Jones}, \footurl{https://lists.w3.org/Archives/Public/semantic-web/2023Sep/0015.html}{Henry Story}, \footurl{https://staffnet.cs.manchester.ac.uk/newsletters/CS_Newsletter_2014-03-31.htm}{Chris Connolly}, \footurl{https://mancunion.com/2022/12/13/make-sure-it-never-happens-again-students-gather-in-memory-of-dr-lloyd-cawthorne/}{Lloyd Cawthorne}, and other such brilliant and lovely people. 

In memoriam: Damon Harvey, Luke O'Connor, Cian Chantrill, Brianna Ghey, Ben Trueman, Ryan Watson, Finn Kitson, Rory Wood, William King, Harrison De George, Charlotte Burlace-Colquhoun, Finn Kitson, Charley Gadd, Sumanta Bansi, Rhys Hill, Laura Nuttall and other students whose lives were cut too short.

My thoughts are with the victims of war, conflict, crime and abuse in Ukraine, Palestine, Israel, Yemen, Sudan, Mexico and across the world.

\newpage
\section{My funding}

My work presented in this thesis has been undertaken during 
several research projects at The University of Manchester, 
which funding is acknowledged below. I am grateful for all the discussions, collaborations and in these projects.

\begin{longtable}{rlll}
\multicolumn{3}{l}{\textbf{European Commission programme Horizon H2020}} 
\\
H2020-INFRAEDI-02-2018 & 
\footurl{https://doi.org/10.3030/823830}{823830} & 
\footurl{https://bioexcel.eu/}{BioExcel-2} & 
\\
H2020-INFRAEOSC-2018-2 & 
\footurl{https://doi.org/10.3030/824087}{824087} & 
\footurl{https://www.eosc-life.eu/}{EOSC-Life} & 
\\
H2020-INFRAIA-2017-1-two-stage & 
\footurl{https://doi.org/10.3030/730976}{730976} & 
\footurl{https://ibisba.eu/}{IBISBA~1.0}
\\
H2020-INFRAIA-2018-1 & 
\footurl{https://doi.org/10.3030/823827}{823827} & 
\footurl{https://www.synthesys.info/}{SyntheSys+}
\\
\\
\multicolumn{3}{l}{\textbf{European Commission programme Horizon Europe}} & \textbf{UKRI}\footnote{UK Research and Innovation under the UK government’s  \emph{Horizon Europe funding guarantee}} 
\\
HORIZON-INFRA-2021-EMERGENCY-01 & 
\footurl{https://doi.org/10.3030/101046203}{101046203} & 
\footurl{https://by-covid.eu/}{BY-COVID}
\\
HORIZON-INFRA-2021-EOSC-01 & 
\footurl{https://doi.org/10.3030/101057388}{101057388} & 
\footurl{http://eurosciencegateway.eu/}{EuroScienceGateway} & 
\footurl{https://gtr.ukri.org/projects?ref=10038963}{10038963}
\\
HORIZON-INFRA-2021-TECH-01 & 
\footurl{https://doi.org/10.3030/101057437}{101057437} & 
\footurl{https://biodt.eu/}{BioDT} & 
\footurl{https://gtr.ukri.org/projects?ref=10038930}{10038930}
\\
HORIZON-INFRA-2021-EOSC-01-05 & 
\footurl{https://doi.org/10.3030/101057344}{101057344} & 
\footurl{https://fair-impact.eu/}{FAIR-IMPACT} & 
\footurl{https://gtr.ukri.org/projects?ref=10038992}{10038992} 
\\

\\
\multicolumn{3}{l}{\textbf{UK Research and Innovation (UKRI)}} \\
MRC / \footurl{https://dareuk.org.uk/driver-project-tre-fx/}{DARE-UK}\footnote{Call: \textit{Inform design of cross-council trusted research environments}} & 
	&
	\footurl{https://trefx.uk/}{TRE-FX} &
\footurl{https://gtr.ukri.org/projects?ref=MC_PC_23007}{MC\_PC\_23007}  

\end{longtable}

\section{Funding and acknowledgements for co-authored chapters}

%%

\subsection{Acknowledgements for \textit{Evaluating FAIR Digital Object and Linked Data as distributed object systems}}\label{ch11:fdo}

Sections \vref{ch3:background} and \vref{ch3:evaluating-fdo-ld} are adapted from a journal article accepted for publication in PeerJ Computer Science.

\subsubsection*{Published As}

Stian Soiland-Reyes, Carole Goble, Paul Groth (2023):\\
\textbf{Evaluating FAIR Digital Object and Linked Data as distributed object systems}.\\
\emph{PeerJ Computer Science} (accepted) \\
\emph{arXiv} 2306.07436 [cs.DC] \\
\url{https://doi.org/10.48550/arXiv.2306.07436}

An \footurl{https://w3id.org/ro/doi/10.5281/zenodo.8075229}{RO-Crate for this article} is archived in Zenodo \cite{Soiland-Reyes 2023a}.


\subsubsection*{Acknowledgements}

We would like to acknowledge the \footurl{https://fairdo.org/}{FAIR Digital Object Forum} community and working groups, where SSR and CG are members.

Views and opinions expressed in this work are those of the authors only and do not necessarily reflect those of the funded projects, FAIR Digital Object Forum, European Union nor the European Commission.

\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  \href{https://spdx.org/licenses/CC-BY-4.0}{(CC BY 4.0)}.
\item
  \textbf{Modifications}: Minor LaTeX changes; references in s11 house style; 
  citations merged and renumbered; 
  funding and references moved to separate chapters. Background moved to separate chapter, added footnote to define \emph{machine actionable}, added figures \ref{ch3:fig:fdo}, \ref{ch3:fig:jsonld} and listing \ref{ch3:triples}. Minor typos and grammatical errors fixed.
\end{itemize}

\subsubsection*{Funding}

This work was funded by the European Union programmes \emph{Horizon 2020} under grant agreements H2020-INFRAEDI-02-2018 823830 (BioExcel-2), H2020-INFRAEOSC-2018-2 824087 (EOSC-Life) and \emph{Horizon Europe} under grant agreements HORIZON-INFRA-2021-EMERGENCY-01 101046203 (BY-COVID), HORIZON-INFRA-2021-EOSC-01 101057388 (EuroScienceGateway), HORIZON-INFRA-2021-EOSC-01-05 101057344 (FAIR-IMPACT), HORIZON-INFRA-2021-TECH-01 101057437 (BioDT), HORIZON-CL4-2021-HUMAN-01-01 101070305 (ENEXA); and by UK Research and Innovation (UKRI) under the UK government’s \emph{Horizon Europe funding guarantee} grants 10038963 (EuroScienceGateway), 10038992(FAIR-IMPACT), 10038930 (BioDT).

Views and opinions expressed in this work are those of the authors only and do not necessarily reflect those of the funded projects, FAIR Digital Object Forum, European Union nor the European Commission.


%%

\subsection{Acknowledgements for \textit{Updating Linked Data practices for FAIR Digital Object principles}}\label{ch11:updating-ld}


Section \ref{ch2:updating-linked-data-practices-for-fair-digital-object-principles} is adapted from a published peer-reviewed conference abstract,
presented as talk by Stian Soiland-Reyes at 
First International Conference on FAIR Digital Objects\footnote{\url{https://www.fdo2022.org/}} 
(FDO2022) on
2022-08-26/--28 in Leiden, The Netherlands. 


\begin{itemize}
    \item Slides: \url{https://doi.org/10.5281/zenodo.7256428}
\end{itemize}

\subsubsection*{Published As}

Stian Soiland-Reyes, Leyla Jael Castro, Daniel Garijo, Marc Portier, Carole Goble, Paul Groth (2022):\\
\textbf{Updating Linked Data practices for FAIR Digital Object principles}.\\
1st International Conference on FAIR Digital Objects (FDO 2022) (abstract).\\
\textit{Research Ideas and Outcomes} \textbf{8}:e94501\\
\url{https://doi.org/10.3897/rio.8.e94501}


\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; figure caption
		formatting; reference in s11 house style\footnote{\url{https://s11.no/2021/house-rules/citation-style/}}; citations merged and renumbered; new introduction;
  acknowledgements and references moved to separate chapters, fixed minor typos and grammatical errors.
\end{itemize}

\subsubsection*{Acknowledgements}

We would like to acknowledge the
\footurl{https://www.researchobject.org/ro-crate/community.html}{RO-Crate
community} and the
\footurl{https://about.workflowhub.eu/project/acknowledgements/}{WorkflowHub
Club}. Thanks to Rudolf Wittner for valuable comments.

\subsubsection*{Funding}

European Commission Horizon 2020 (EOSC-Life
\footurl{https://doi.org/10.3030/824087}{824087}), Horizon
Europe (BY-COVID
\footurl{https://doi.org/10.3030/101046203}{101046203},
FAIR-IMPACT
\footurl{https://doi.org/10.3030/101057344}{101057344}).

Leyla Jael Castro is supported by a German Research Foundation DFG grant
for NFDI4DataScience.

Daniel Garijo is supported by the Madrid Government (Comunidad de
Madrid-Spain) under the Multiannual Agreement with Universidad
Politécnica de Madrid in the line Support for R\&D projects for Beatriz
Galindo researchers, in the context of the V PRICIT (Regional Programme
of Research and Technological Innovation)



%% 

\subsection{Acknowledgements for \emph{Packaging research artefacts with RO-Crate}}\label{ch11:packagingrocrate}

Sections \vref{ch5:packaging-research-artefacts-with-ro-crate} and \vref{ch5:formaldefinition} are adapted from an article published in the journal \emph{Data Science}.

\subsubsection*{Published As}

Stian Soiland-Reyes, Peter Sefton, Mercè Crosas, Leyla Jael Castro,
Frederik Coppens, José M. Fernández, Daniel Garijo, Björn Grüning, Marco
La Rosa, Simone Leo, Eoghan Ó Carragáin, Marc Portier, Ana Trisovic,
RO-Crate Community, Paul Groth, Carole Goble (2022):\\
\textbf{Packaging research artefacts with RO-Crate}.\\
\emph{Data Science} \textbf{5}(2)\\
\url{https://doi.org/10.3233/DS-210053}

An \footurl{https://w3id.org/ro/doi/10.5281/zenodo.5146227}{RO-Crate for this article} is archived in Zenodo \cite{Soiland-Reyes 2022g}.

\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; figure caption
  formatting; reference in s11 house style; added identifiers, authors
  and years clarified where missing in citations; inline citation
  hyperlinks to open access version where available; citations merged and renumbered; 
  acknowledgements and references moved to separate chapters; fixed minor typos and grammatical errors.
\end{itemize}


\subsubsection*{Funding}

This work has received funding from the European Commission's Horizon
2020 research and innovation programme for projects
\footurl{https://doi.org/10.3030//823830}{BioExcel-2}
(H2020-INFRAEDI-2018-1 823830),
\footurl{https://doi.org/10.3030/730976}{IBISBA~1.0}
(H2020-INFRAIA-2017-1-two-stage 730976),
\footurl{https://doi.org/10.3030/871118}{PREP-IBISBA}
(H2020-INFRADEV-2019-2 871118),
\footurl{https://doi.org/10.3030/824087}{EOSC-Life}
(H2020-INFRAEOSC-2018-2 824087),
\footurl{https://doi.org/10.3030/823827}{SyntheSys+}
(H2020-INFRAIA-2018-1 823827). From the Horizon Europe Framework
Programme this work has received funding for
\footurl{https://doi.org/10.3030/101046203}{BY-COVID}
(HORIZON-INFRA-2021-EMERGENCY-01 101046203).

Björn Grüning is supported by DataPLANT
\footurl{https://gepris.dfg.de/gepris/projekt/442077441}{(NFDI 7/1 --
42077441)}, part of the German National Research Data Infrastructure
(NFDI), funded by the Deutsche Forschungsgemeinschaft (DFG).

Ana Trisovic is funded by the Alfred P. Sloan Foundation.
\footurl{https://sloan.org/grant-detail/9555}{(Grant number P-2020-13988)}
Harvard Data Commons is supported by an award from Harvard University
Information Technology (HUIT).


\subsubsection*{Contributions}

We would also like to acknowledge contributions from:

\begin{description}
\tightlist
\item[Finn Bacall]
Software, Methodology
\item[Herbert Van de Sompel]
Writing -- review \& editing
\item[Ignacio Eguinoa]
Software, Methodology
\item[Nick Juty]
Writing -- review \& editing
\item[Oscar Corcho]
Writing -- review \& editing
\item[Stuart Owen]
Writing -- review \& editing
\item[Laura Rodríguez-Navas]
Software, Visualization, Writing -- review \& editing
\item[Alan R. Williams]
Writing -- review \& editing
\end{description}


%%


\subsection{Acknowledgements for \textit{Creating lightweight FAIR Digital Objects with RO-Crate}}\label{ch11:lightweight}

Section \vref{ch4:lightweight-fdo} is adapted from a published peer-reviewed conference abstract,
presented as poster by Stian Soiland-Reyes at 
First International Conference on FAIR Digital Objects 
\footurl{https://www.fdo2022.org/}{(FDO2022)} on
022-08-26/--28 in Leiden, The Netherlands. 

\begin{itemize}
\item
  Poster: \url{https://doi.org/10.5281/zenodo.7245315}
\end{itemize}

\subsubsection*{Published As}
Stian Soiland-Reyes, Peter Sefton, Leyla Jael Castro, Frederik Coppens,
Daniel Garijo, Simone Leo, Marc Portier, Paul Groth (2022):\\
\textbf{Creating lightweight FAIR Digital Objects with RO-Crate}.\\
1st International Conference on FAIR Digital Objects (FDO2022) (poster)\\
\emph{Research Ideas and Outcomes} \textbf{8}:e93937\\
\url{https://doi.org/10.3897/rio.8.e93937}


\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; figure caption
  formatting; references in s11 house style; citations merged and renumbered; 
  acknowledgements and references moved to separate chapters. 
  Figure \ref{fig:signposting} font re-rendered. Added figure \ref{ch9:fig:signposting}. Fixed minor typos and grammatical errors
\end{itemize}

\subsubsection*{Acknowledgements}

We would like to acknowledge the
\footurl{https://www.researchobject.org/ro-crate/community.html}{RO-Crate
community} and the
\footurl{https://about.workflowhub.eu/project/acknowledgements/}{WorkflowHub
Club}.

\subsubsection*{Funding}

European Commission Horizon 2020 (BioExcel-2
\footurl{https://doi.org/10.3030/823830}{823830}, EOSC-Life
\footurl{https://doi.org/10.3030/824087}{824087}), Horizon
Europe (BY-COVID
\footurl{https://doi.org/10.3030/101046203}{101046203},
FAIR-IMPACT
\footurl{https://doi.org/10.3030/101057344}{101057344}).

Daniel Garijo is supported by the Madrid Government (Comunidad de
Madrid-Spain) under the Multiannual Agreement with Universidad
Politécnica de Madrid in the line Support for R\&D projects for Beatriz
Galindo researchers, in the context of the V PRICIT (Regional Programme
of Research and Technological Innovation).

Leyla Jael Castro is supported by a German Research Foundation DFG grant
for NFDI4DataScience.

Frederik Coppens is supported by Research Foundation - Flanders (FWO)
for ELIXIR Belgium (I002819N).


%%

\subsection{Acknowledgements for \emph{Making Canonical Workflow Building Blocks}}\label{ch11:canonical}

Section \vref{ch6:making-canonical-workflow-building-blocks-interoperable-across-workflow-languages} is adapted from journal article published in \emph{Data Intelligence}.

\subsubsection*{Published As}

Stian Soiland-Reyes, Genís Bayarri, Pau Andrio, Robin Long, Douglas
Lowe, Ania Niewielska, Adam Hospital, Paul Groth (2022):\\
\textbf{Making Canonical Workflow Building Blocks interoperable across
workflow languages}.\\
\emph{Data Intelligence} \textbf{4}(2)\\
\url{https://doi.org/10.1162/dint_a_00135}


\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; figure caption
  formatting; references in s11 house style; citations merged and renumbered; 
  acknowledgements and references moved to separate chapters. 
  Fixed minor typos and grammatical errors.
\end{itemize}



\subsubsection*{Acknowledgements}

This work has been done as part of the BioExcel CoE
(\url{https://www.bioexcel.eu/}), a project funded by the European Union
contracts
\footurl{https://doi.org/10.3030/823830}{H2020-INFRAEDI-02-2018
823830},
\footurl{https://doi.org/10.3030/675728}{H2020-EINFRA-2015-1
675728}. Additional work is funded through EOSC-Life
(\url{https://www.eosc-life.eu/}) contract
\footurl{https://doi.org/10.3030/824087}{H2020-INFRAEOSC-2018-2
824087}, and ELIXIR-CONVERGE (\url{https://elixir-europe.org/}) contract
\footurl{https://doi.org/10.3030/871075}{H2020-INFRADEV-2019-2
871075}.

The authors would also like to acknowledge contributions from: Felix
Amaladoss, Cibin Sadasivan Baby, Finn Bacall, Rosa M. Badia, Sarah
Butcher, Gerard Capes, Michael R. Crusoe, Alberto Eusebi, Carole Goble,
Josep Lluís Gelpí, Modesto Orozco, Geoff Williams, Felix Amaladoss

%% 

\subsection{Acknowledgements for \emph{The Specimen Data Refinery}}\label{ch11:refinery}

Section \vref{ch8:the-specimen-data-refinery} is adapted from a journal article published in \emph{Data Intelligence}.

\subsubsection*{Published As}

Alex Hardisty, Paul Brack, Carole Goble, Laurence Livermore, Ben Scott,
Quentin Groom, Stuart Owen, Stian Soiland-Reyes (2022):\\
\textbf{The Specimen Data Refinery: A canonical workflow framework and
FAIR Digital Object approach to speeding up digital mobilisation of
natural history collections.}\\
\emph{Data Intelligence} \textbf{4}(2)\\
\url{https://doi.org/10.1162/dint_a_00134}

\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; figures replaced with higher resolutions from source; figure caption formatting; references in s11 house style; added DOIs and URLs; cited preprints replaced with later publications
  citations merged and renumbered; funding and references moved to separate chapters. Fixed minor typos and grammatical errors.
\end{itemize}


\subsubsection*{Acknowledgements}

This work has received funding from the European Union's Horizon 2020
research and innovation programme under grant agreement numbers 823827
(SYNTHESYS Plus), 871043 (DiSSCo Prepare), 823830 (BioExcel-2), 824087
(EOSC-Life).



%% 

\subsection{Acknowledgements for \emph{Incrementally building FAIR Digital Objects}}\label{ch11:incrementally-fdo}

Section \vref{ch7:incrementally-building-fair-digital-objects-with-specimen-data-refinery-workflows}  is adapted from an abstract
presented as poster by Stian Soiland-Reyes at 
First International Conference on FAIR Digital Objects\footnote{\url{https://www.fdo2022.org/}}
(FDO2022) on
2022-08-26/--28 in Leiden, The Netherlands. 

\subsubsection*{Published As}

Oliver Woolland, Paul Brack, Stian Soiland-Reyes, Ben Scott, Laurence
Livermore (2022):\\
\textbf{Incrementally building FAIR Digital Objects with Specimen Data
Refinery workflows}.\\
1st International Conference on FAIR Digital Objects
(FDO 2022) (poster)\\
\emph{Research Ideas and Outcomes} \textbf{8}:e94349\\
\url{https://doi.org/10.3897/rio.8.e94349}



\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and LaTeX; references in s11 house style; added DOIs and URLs; funding and references moved to separate chapters; fixed minor typos and grammatical errors.
\end{itemize}


\subsubsection*{Acknowledgements}

We acknowledge the \footurl{https://www.synthesys.info/}{SYNTHESYS+} and
\footurl{https://www.dissco.eu/}{DiSSCO} project members who have been
invaluable in early evaluation and feedback on the development of SDR.

\subsubsection*{Funding}

This work has received funding from the European Union's Horizon 2020
research and innovation programme under grant agreement numbers
\footurl{https://doi.org/10.3030/823827}{823827}
(SYNTHESYS Plus), \footurl{https://doi.org/10.3030/871043}{871043} (DiSSCo
Prepare), \footurl{https://doi.org/10.3030/823830}{823830} (BioExcel-2),
\footurl{https://doi.org/10.3030/824087}{824087} (EOSC-Life).



\subsection{Acknowledgement for \emph{Recording provenance of workflow runs with RO-Crate}}
\label{ch11:wrroc}

Section \vref{ch54:wrroc} is adapted from an arXiv preprint submitted to PLOS One.

\subsubsection*{Published As}

Simone Leo, Michael R. Crusoe, Laura Rodríguez-Navas, Raül Sirvent, Alexander Kanitz, Paul De Geest, Rudolf Wittner, Luca Pireddu, Daniel Garijo, José M. Fernández, Iacopo Colonnelli, Matej Gallo, Tazro Ohta, Hirotaka Suetake, Salvador Capella-Gutierrez, Renske de Wit, Bruno de Paula Kinoshita, Stian Soiland-Reyes (2023): \\
\textbf{Recording provenance of workflow runs with RO-Crate}.\\
\emph{arXiv} 2312.07852 [cs.DL] \\
\url{https://doi.org/10.48550/arXiv.2312.07852}

An \footurl{https://w3id.org/ro/doi/10.5281/zenodo.10368989}{RO-Crate for this article} is archived in Zenodo \cite{Leo 2023c}.


\subsubsection*{License and modifications}

\begin{itemize}
\tightlist
\item
  \textbf{License}: Creative Commons Attribution License
  (\href{https://spdx.org/licenses/CC-BY-4.0}{CC BY 4.0}).
\item
  \textbf{Modifications}: Formatting as Markdown and figure caption formatting; references in s11 house style; URLs as footnotes/hyperlinks; enumerations made explicit; listing captions; some paragraphs split for readability; details moved to footnote, acknowledgement and references moved to separate chapters; fixed minor typos and grammatical errors.
\end{itemize}


\subsubsection*{Acknowledgements}

The authors would like to thank all participants to the 
\footurl{https://www.researchobject.org/workflow-run-crate/\#community}{Workflow Run RO-Crate working group} 
meetings for the positive detailed discussions and
valuable feedback.

The authors acknowledge funding from:
  Sardinian Regional Government through the XData Project (S.L., L.P.);
  Spanish Government (contract PID2019-107255GB) (R.S.);
  MCIN/AEI/10.13039/501100011033 (CEX2021-001148-S) (R.S.);
  Generalitat de Catalunya (contract 2021-SGR-00412) (R.S.);
  European High-Performance Computing Joint Undertaking (JU) (No 
  \footurl{https://doi.org/10.3030/955558}{955558}) (R.S.);
  EU Horizon research and innovation programme under Grant agreement No 
  \footurl{https://doi.org/10.3030/101058129}{101058129} (DT-GEO) (R.S.);
  ELIXIR Platform Task 2022-2023 funding for Task ``Container Orchestration'' (A.K.);
  Research Foundation - Flanders (FWO) for ELIXIR Belgium (I000323N and I002819N) (P.D.G.);
  Multiannual Agreement with Universidad Politécnica de Madrid in the line Support for R\&D projects for Beatriz Galindo researchers, in the context of the V PRICIT (Regional Programme of Research and Technological Innovation) (D.G.);
  Comunidad de Madrid through the call Research Grants for Young Investigators from Universidad Politécnica de Madrid (D.G.);
  ICSC - Centro Nazionale di Ricerca in High-Performance Computing, Big Data and Quantum Computing, funded by European Union - NextGenerationEU (I.C.);
  ACROSS project, HPC Big Data Artificial Intelligence Cross Stack Platform Towards Exascale, funded by the European High-Performance Computing Joint Undertaking (JU) under G.A. n. 
  \footurl{https://doi.org/10.3030/955648}{955648} (I.C.);
  EUPEX project, European Pilot for Exascale, funded by the European High-Performance Computing Joint Undertaking (JU) under G.A. n. 
  \footurl{https://doi.org/10.3030/101033975}{101033975} (I.C.);
  Life Science Database Integration Project, NBDC of Japan Science and
  Technology Agency (T.O.);
  JSPS KAKENHI (Grant Number \footurl{https://kaken.nii.ac.jp/en/grant/KAKENHI-PROJECT-20J22439/}{20J22439});
  European Commission Horizon 2020
  H2020-SC1-2018-Single-Stage-RTD
\footurl{https://doi.org/10.3030/825575}{825575} (European Joint Programme on Rare Diseases; SC1-BHC-04-2018 Rare Disease European Joint Programme Cofund) (L.R.N., J.M.F., S.C.G.),
    European High-Performance Computing Joint Undertaking (JU) (No 955558),
  EU NextGenerationEU/PRTR (project eFlows4HPC)
  H2020-JTI-EuroHPC-2019-1 
\footurl{https://doi.org/10.3030/955558}{955558} (eFlows4HPC) (R.S.),
  H2020-INFRAEDI-02-2018 
\footurl{https://doi.org/10.3030/823830}{823830}
  (BioExcel-2) (S.S.R.),
  H2020-INFRAEOSC-2018-2 
\footurl{https://doi.org/10.3030/824087}{824087}
  (EOSC-Life) (S.L., L.R.N., P.D.G., R.W., L.P., J.M.F., S.C.G., S.S.R.);
  Horizon Europe
  HORIZON-INFRA-2021-EMERGENCY-01
\footurl{https://doi.org/10.3030/101046203}{101046203} (BY-COVID) (S.L., L.R.N., P.D.G., R.W., L.P., J.M.F., S.C.G., S.S.R.),
  HORIZON-INFRA-2021-EOSC-01
\footurl{https://doi.org/10.3030/101057388}{101057388} (EuroScienceGateway) (P.D.G., J.M.F., S.C.G., S.S.R.),
  HORIZON-INFRA-2021-EOSC-01-05
\footurl{https://doi.org/10.3030/101057344}{101057344} (FAIR-IMPACT) (D.G., S.S.R.);
  UK Research and Innovation (UKRI) under the UK government's Horizon
  Europe funding guarantee
\footurl{https://gtr.ukri.org/projects?ref=10038963}{10038963} (EuroScienceGateway),
\footurl{https://gtr.ukri.org/projects?ref=10038992}{10038992} (FAIR-IMPACT) (S.S.R.).

