%% Unmerged references
%%%%
%%% Below is NOT MERGED
%%% 

\section{References from Chapter 3}

\begin{thebibliography}{9}
   
\bibitem[Wilkinson 2016]{monsCloudyIncreasinglyFAIR2017b}
Mark D. Wilkinson, Michel Dumontier, IJsbrand Jan Aalbersberg, Gabrielle
Appleton, Myles Axton, Arie Baak, Niklas Blomberg, Jan-Willem Boiten,
Luiz Bonino da Silva Santos, Philip E. Bourne, Jildau Bouwman, Anthony
J. Brookes, Tim Clark, Mercè Crosas, Ingrid Dillo, Olivier Dumon, Scott
Edmunds, Chris T. Evelo, Richard Finkers, Alejandra Gonzalez-Beltran,
Alasdair J. G. Gray, Paul Groth, Carole Goble, Jeffrey S. Grethe, Jaap
Heringa, Peter A. C. 't Hoen, Rob Hooft, Tobias Kuhn, Ruben Kok, Joost
Kok, Scott J. Lusher, Maryann E. Martone, Albert Mons, Abel L. Packer,
Bengt Persson, Philippe Rocca-Serra, Marco Roos, Rene van Schaik,
Susanna-Assunta Sansone, Erik Schultes, Thierry Sengstag, Ted Slater,
George Strawn, Morris A. Swertz, Mark Thompson, Johan van der Lei, Erik
van Mulligen, Jan Velterop, Andra Waagmeester, Peter Wittenburg,
Katherine Wolstencroft, Jun Zhao, Barend Mons (2016):\\
The {FAIR
Guiding Principles} for scientific data management and stewardship.
\emph{Scientific Data} \textbf{3} \\
\url{https://doi.org/10.1038/sdata.2016.18}

\bibitem[Mons 2017]{monsCloudyIncreasinglyFAIR2017b}
Barend Mons, Cameron Neylon, Jan Velterop, Michel Dumontier, Luiz Olavo
Bonino Silva Santos, Mark D. Wilkinson (2017): \\
\textbf{Cloudy, increasingly {FAIR}; revisiting the {FAIR} data guiding principles for the European
Open Science Cloud}.\\
\emph{Information Services \& Use} \textbf{37}(1)\\
\url{https://doi.org/10.3233/ISU-170824}

\bibitem[Corcho 2021]{corchoEOSCInteroperabilityFramework2021b}
Oscar Corcho, Magnus Eriksson, Krzysztof Kurowski, Milan Ojsteršek,
Christine Choirat, Mark Sanden, Frederik Coppens, EOSC Executive
Board (2021): \\
\textbf{EOSC Interoperability Framework}.\\
\emph{Publications Office of the EU} \\
\url{https://doi.org/10.2777/620649}

\bibitem[Schultes 2019]{schultesFAIRPrinciplesDigital2019a}
Erik Schultes, Peter Wittenburg (2019): \\
\textbf{FAIR principles and digital objects: Accelerating convergence on a data infrastructure}.\\
\emph{Data analytics and management in data intensive domains}: 20th international conference,
{DAMDID}/{RCDL} 2018, Moscow, Russia, 2018-10-09/--12. \\
\url{https://doi.org/10.1007/978-3-030-23584-0_1}\\
Preprint: \url{https://doi.org/10.23728/B2SHARE.166A074BFF614A31B05E9DF5BFD9809D}

\bibitem[Bizer 2009]{bizerLinkedDataStory2009a}
Christian Bizer, Tom Heath, Tim Berners-Lee (2009): \\
\textbf{Linked data - the story so far}. \\
\emph{International journal on Semantic Web and information systems} \textbf{5} \\
\url{https://doi.org/10.4018/jswis.2009081901}

\bibitem[Bonino Da Silva Santos 2016]{boninodasilvasantosFAIRDataPoints2016a}
Luiz Olavo Bonino Da Silva Santos, Mark D. Wilkinson, Arnold Kuzniar,
Rajaram Kaliyaperumal, Mark Thompson, Michel Dumontier, Kees Burger
(2016):\\
\textbf{FAIR Data points supporting big data interoperability}. \\
\emph{Enterprise interoperability in the digitized and networked factory of the future}, Martin Zelm, Guy Doumeingts, Joao Pedro Mendonça (eds.).\\
iSTE Press. \\
ISBN 978-1-84704-044-2 \\
\url{https://www.researchgate.net/publication/309468587_FAIR_Data_Points_Supporting_Big_Data_Interoperability}

\bibitem[Hasnain 2018]{hasnainAssessingFAIRData2018a}
Ali Hasnain, Dietrich Rebholz-Schuhmann (2018): \\
\textbf{Assessing FAIR data principles against the 5-star open data principles}. \\
\emph{The semantic web: ESWC 2018 satellite events}, Heraklion, Crete, Greece, 2018-06-03/--07 \\
\url{https://doi.org/10.1007/978-3-319-98192-5_60}

\bibitem[Verborgh 2020]{verborghSemanticWebIdentity2020a}
Ruben Verborgh, Miel Vander Sande (2020): \\
\textbf{The semantic web identity crisis: In search of the trivialities that never were}.\\
\emph{Semantic Web}, \textbf{11}(1)\\
\url{https://doi.org/10.3233/SW-190372}

\bibitem[Meroño-Peñuela 2021]{merono-penuelaConclusionFutureChallenges2021a}
Albert Meroño-Peñuela, Pasquale Lisena, Carlos Martínez-Ortiz (2021):
\textbf{Conclusion and future challenges}. \emph{Web data apis for
knowledge graphs: {Easing} access to semantic data for application
developers} ({Cham}: {Springer International Publishing}, 2021), pp.
77--80.
\url{https://doi.org/10.1007/978-3-031-01917-3_7}

\bibitem[Neumann 2021]{neumannAnalysisPublicREST2021a}
Andy Neumann, Nuno Laranjeiro, Jorge Bernardino (2021): \\
\textbf{An analysis of public {REST} web service apis}.\\
\emph{IEEE Transactions on Services Computing} \textbf{14}
\url{https://doi.org/10.1109/TSC.2018.2847344}

\bibitem[Turcoane 2014]{turcoaneLinkedDataJSONLD2014a}
Ovidiu Turcoane (2014): \\
\textbf{Linked data, {JSON-LD} and the semantics of cultural and scientific heritage}. \\
\emph{Digital Presentation and Preservation of Cultural and Scientific Heritage} \textbf{4} \\
\url{https://doi.org/10.55630/dipp.2014.4.11}

\bibitem[Wilkinson 2022]{wilkinsonWorkflowsWhenParts2022b}
Sean R. Wilkinson, Greg Eisenhauer, Anuj J. Kapadia, Kathryn Knight,
Jeremy Logan, Patrick Widener, Matthew Wolf (2022): \\
\textbf{F*** workflows: When parts of {FAIR} are missing}. 
\emph{arXiv} 2209.09022 [cs.DL] \\
\url{https://doi.org/10.48550/arxiv.2209.09022}

\bibitem[Stefi 2015]{stefiDevelopersMakeUnbiased2015}
Anisa Stefi (2015): \\
\textbf{Do Developers Make Unbiased Decisions? - The Effect of Mindfulness and Not-Invented-Here Bias on the Adoption of Software Components}. \\
\emph{European Conference on Information Systems} (ECIS 2015), Münster, Germany, 2015-05-26/--29 \\
\url{https://doi.org/10.18151/7217489}

\bibitem[Stefi 2015]{stefiDevelopReuseTwo2015a}
Anisa Stefi, Thomas Hess (2015): \\
\textbf{To develop or to reuse? Two perspectives on external reuse in software projects}. \\
\emph{International Conference of Software Business} (ICSOB 2015), Braga, Portugal, 2015-06-10/--12.\\
\emph{ICSOB 2015: Software business} \\
\url{http://dx.doi.org/10.1007/978-3-319-19593-3_18}

\bibitem[Meroño-Peñuela 2021]{merono-penuelaWebDataApis2021b}
Albert Meroño-Peñuela, Pasquale Lisena, Carlos Martínez-Ortiz (2021):
\textbf{Web data APIs over {SPARQL}}. \\
in: \emph{Web Data APIs for Knowledge Graphs : Easing access to semantic data for application developers}, 
Albert Meroño-Peñuela, Pasquale Lisena, Carlos Martínez-Ortiz.
Synthesis Lectures on Data, Semantics, and Knowledge.
\url{https://doi.org/10.1007/978-3-031-01917-3_3}

\bibitem[Kahn 2006]{kahnFrameworkDistributedDigital2006b}
Robert Kahn, Robert Wilensky (2006): \\
\textbf{A framework for distributed digital object services}.\\
\emph{International Journal on Digital Libraries} \textbf{6} \\
\url{https://doi.org/10.1007/s00799-005-0128-x}

\bibitem[Kahn 1995]{kahnFrameworkDistributedDigital1995a}
Robert Kahn, Robert Wilensky (1995): \\
\emph{A framework for distributed digital object services} (CNRI).
\url{http://www.cnri.reston.va.us/k-w.html} (accessed 9 May 2022)

\bibitem[ITU-T X.1255]{x1255FrameworkDiscovery}
ITU-T (2013): \\
\textbf{X.1255 : Framework for Discovery of Identity Management Information}.\\
\emph{Series X: Data networks, open system communications and security} ITU-T X.1255 \\
The International Telecommunication Union (ITU).\\
\url{https://www.itu.int/rec/T-REC-X.1255-201309-I}

\bibitem[Reilly 2009]{DigitalObjectInterface}
Sean Reilly (2009): \\
\textbf{Digital Object Interface Protocol Version 1.0}.
(2009). \url{https://www.dona.net/doipv1doc} (accessed 26 May 2022)

\bibitem[Wittenburg 2019]{wittenburgDigitalObjectsDrivers2019a}
Peter Wittenburg, George Strawn, Barend Mons, Luiz Bonino, Erik
Schultes (2019): \\
Digital objects as drivers towards convergence in data
infrastructures. \emph{https://b2share.eudat.eu}, (2019).
\url{https://doi.org/10.23728/b2share.b605d85809ca45679b110719b6c6cb11}

\bibitem[Bonino 2019]{boninoFAIRDigitalObject}
Luiz Bonino, Oeter Wittenburg, Bonnie Carroll, Alex Hardisty, Mark
Leggott, Carlo Zwölf (2019): \\
\textbf{{FAIR} digital object framework}. \\
\url{https://github.com/GEDE-RDA-Europe/GEDE/blob/master/FAIR\%20Digital\%20Objects/FDOF/FAIR\%20Digital\%20Object\%20Framework-v1-02.docx}

\bibitem[Wittenburg 2022]{wittenburgFAIRDigitalObject2022b}
Peter Wittenburg, Ivonne Anders, Christophe Blanchi, Merret Buurman,
Carole Goble, Jonas Grieb, Alex Hardisty, Sharif Islam, Thomas Jejkal,
Tibor Kálmán, Christine Kirkpatrick, Laurence Lannom, Thomas Lauer,
Giridhar Manepalli, Karsten Peters-von Gehlen, Andreas Pfeil, Robert
Quick, Mark Sanden, Ulrich Schwardmann, Stian Soiland-Reyes, Rainer
Stotzka, Zachary Trautt, Dieter Van Uytvanck, Claus Weiland, Philipp
Wieder (2022): \\
\textbf{FAIR digital object demonstrators 2021}.\\
\emph{Zenodo},
\url{https://doi.org/10.5281/zenodo.5872645}

\bibitem[DONA 2018]{foundationDigitalObjectInterface}
DONA (2018): \\
\textbf{Digital object interface protocol specification, version 2.0} 
\emph{DONA Foundation}.\\
\url{https://hdl.handle.net/0.DOIP/DOIPV2.0}

\bibitem[Cordra DOIP]{DOIPExamplesCordraa}
(n.d.): {DOIP} and {Examples} --- {Cordra} documentation. (n.d.).
\url{https://www.cordra.org/documentation/api/doip.html} (accessed 26
May 2022)

\bibitem[Wright 2022]{Draftbhuttonjsonschema}
Austin Wright, Henry Andrews, Ben Hutton, Greg Dennis (2022):\\
\textbf{JSON schema: A media type for describing JSON documents}.\\
\emph{Internet Engineering Task Force}\\
\url{https://datatracker.ietf.org/doc/draft-bhutton-json-schema/01/}

\bibitem[FDO]{FAIRDigitalObjects}
FDO: {FAIR Digital Objects Forum} \\
\url{https://fairdo.org/} (accessed 26 May 2022)

\bibitem[FDO Specs]{fdo-Specs}
(2022): \\
\emph{{FDO Specification Documents - November 2022}} ({FDO
Forum}, 2022). \url{https://fairdo.org/specifications/} (accessed 2
February 2023)

\bibitem[FDO Roadmap]{fdo-Roadmap}
Larry Lannom, Ulrich Schwardmann, Christophe Blanchi, Ivonne Anders,
Claus Weiland, Peter Wittenburg (2022): \\
\emph{FAIR digital objects
roadmap. Version 5 november 2022} ({FDO Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7824673}

\bibitem[FDO Overview]{fdo-Overview}
Anders Ivonne, Blanchi Christophe, Broder Daan, Hellström Maggie, Islam
Sharif, Jejkal Thomas, Lannom Larry, Peters-von Gehlen Karsten, Quick
Robert, Schlemmer Alexander, Schwardmann Ulrich, Soiland-Reyes Stian,
Strawn George, Dieter van Uytvanck, Claus Weiland, Peter Wittenburg, \&
Carlo Zwölf (2023): \\
\emph{{FAIR} digital object technical overview.
Version PEN 2.0} ({FAIR Digital Object Forum}, 2023).
\url{https://doi.org/10.5281/zenodo.7824714}

\bibitem[FDO Glossary]{fdo-Glossary}
Daan Broeder, Peter Wittenburg (2022): \\
\emph{{FDO} glossary november
2022} ({FDO Forum}, 2022).
\url{https://drive.google.com/file/d/1KJ9l0p96naKi_2HPJ_MPqPTwS_zlP92G}
(accessed 2 February 2023)

\bibitem[FDO DocProcessStd]{fdo-DocProcessStd}
C. Weiland, U. Schwardmann, P. Wittenburg, C. Kirkpatrick, R. Hanisch,
Z. Trautt, C. Weiland, U. Schwardmann, P. Wittenburg (2022):
\emph{{FDO Forum Document Standards}} ({(internal draft)}: {FDO Forum},
2022).
\url{https://drive.google.com/file/d/1lPNBBROjEoZ6fTfrtdqcMa3Q2G27PoC_}
(accessed 30 November 2022)

\bibitem[FDO RequirementSpec]{fdo-RequirementSpec}
G. Strawn, P. Wittenburg, St. Soiland-Reyes, K. Peters, L. Lannom, U.
Schwardmann, Ch Blanch (2023): \\
\emph{{FDO Forum FDO} requirement
specifications. Version 3.0} ({FDO Forum}, 2023).
\url{https://doi.org/10.5281/zenodo.7782262}

\bibitem[FDO MachineActionDef]{fdo-MachineActionDef}
Claus Weiland, Sharif Islam, Daan Broder, Ivonne Anders, Peter
Wittenburg (2022): \\
\emph{{FDO} machine actionability. Version 2.2} ({FDO
Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825650}

\bibitem[FDO ConfigurationTypes]{fdo-ConfigurationTypes}
Larry Lannom, Karsten Peters-von Gehlen, Ivonne Anders, Andreas Pfeil,
Alexander Schlemmer, Zach Trautt, Peter Wittenburg (2022):
\emph{{FDO} configuration types} ({FAIR Digital Objects Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825703}

\bibitem[Van de Sompel 2022]{vandesompelFAIRSignpostingProfile2022}
Herbert Van de Sompel, Martin Klein, Shawn Jones, Michael L. Nelson,
Simeon Warner, Anusuriya Devaraju, Robert Huber, Wilko Steinhoff,
Vyacheslav Tykhonov, Luc Boruta, Enno Meijers, Stian Soiland-Reyes, \&
Mark Wilkinson (2022): \\
{FAIR Signposting Profile}. (2022).
\url{https://signposting.org/FAIR/} (accessed 5 January 2023)

\bibitem[FDO PIDProfileAttributes]{fdo-PIDProfileAttributes}
Ivonne Anders, Maggie Hellström, Sharif Islam, Thomas Jejkal, Larry
Lannom, Ulrich Schwardmann, Peter Wittenburg (2022): \\
\emph{{FDO PID}
profiles \& attributes} ({FDO Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825630}

\bibitem[weigelRDARecommendationPID2018]{weigelRDARecommendationPID2018}
Tobias Weigel, Beth Plale, Mark Parsons, Gabriel Zhou, Yu Luo, Ulrich
Schwardmann, Robert Quick, Margareta Hellström, Kei Kurakawa (2018):
\emph{{RDA Recommendation} on {PID Kernel Information}} (2018).
\url{https://doi.org/10.15497/rda00031}

\bibitem[islam 2023]{islam_2023}
Sharif Islam (2023): \\
{FAIR} digital objects, persistent identifiers and
machine actionability. \emph{FAIR Connect}, \textbf{1} (2023) 29--34.
\url{https://doi.org/10.3233/FC-230001}

\bibitem[FDO KernelAttributes]{fdo-KernelAttributes}
Daan Broeder, Peter Wittenburg, Ivonne Anders, Karsten Peters-von
Gehlen (2022): \\
\emph{{FDO} -- kernel attributes \& metadata} ({FDO
Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825693}

\bibitem[FDO Granularity]{fdo-Granularity}
Maggie Hellström, Carlo Zwölf, Peter Wittenburg (2022): \\
\emph{{FDO}
-- granularity, versioning, mutability} ({FDO Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825686}

\bibitem[FDO DOIPEndorsement]{fdo-DOIPEndorsement}
Ulrich Schwardmann, George Strawn, Robert Quick, Peter Wittenburg
(2022): \\
\emph{{DOIP} endorsement request} ({FDO Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7824796}

\bibitem[FDO FDO-Upload]{fdo-FDO-Upload}
Christophe Blanchi, Daan Broeder, Thomas Jejkal, Islam Sharif, Alexander
Schlemmer, Dieter van Uytvanck, Peter Wittenburg (2022): \\
\emph{{FDO}
-- upload of {FDO}} ({FAIR Digital Object Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825549}

\bibitem[ResourceSyncFrameworkSpecification]{ResourceSyncFrameworkSpecification}
(2017): \\
{ResourceSync} framework specification ({ANSI/NISO
Z39.99-2017}). (2017).
\url{http://www.openarchives.org/rs/1.1/resourcesync}

\bibitem[FDO TypingFDOs]{fdo-TypingFDOs}
Larry Lannom, Ulrich Schwardmann, Cristophe Blanchi, Peter Wittenburg
(2022): \\
\emph{Typing {FAIR} digital objects} ({FDO Forum}, 2022).
\url{https://doi.org/10.5281/zenodo.7825599}

\bibitem[FDO ImplAttributesTypesProfiles]{fdo-ImplAttributesTypesProfiles}
C. Blanchi, M. Hellström, L. Lannom, U. Schwardmann, P. Wittenburg, U.
Schwardmann, P. Wittenburg (2022): \\
\emph{Implementation of
attributes, types, profiles and registries} ({FAIR Digital Objects
Forum}, 2022).
\url{https://drive.google.com/file/d/1RrOiwMhkl-hRzWmlluA2iXCzHK-bj7_80LlMXgWx4w4}
(accessed 30 November 2022)

\bibitem[Schwardmann 2022]{schwardmannTwoExamplesHow2022}
Ulrich Schwardmann, Tibor Kálmán (2022): \\
Two {Examples} on {How FDO
Types} can {Support Machine} and {Human Readability}. \emph{Research
Ideas and Outcomes}, \textbf{8} (2022).
\url{https://doi.org/10.3897/rio.8.e96014}

\bibitem[Berners-Lee 199]{berners-leeWeavingWebOriginal1999}
Tim Berners-Lee, Mark Fischetti (1999): \\
\textbf{Weaving the {Web}: The original design and ultimate destiny of the {World Wide Web} by its
inventor}.\\
ISBN 978-0-06-251586-5

\bibitem[Schreiber 2014]{w3-rdf11-primer}
Guus Schreiber and Yves Raimond (2014):
\textbf{RDF 1.1 Primer}. \\
\emph{W3C} Note \\
\url{http://www.w3.org/TR/2014/NOTE-rdf11-primer-20140624}

\bibitem[W3 RDF/XML]{w3-rdf-syntax99}
Ora Lassila, Ralph R. Swick (1999): \\
\textbf{Resource {Description Framework} ({RDF}) {Model} and {Syntax Specification}}. \\
\emph{W3C} Recommendation\\
\url{https://www.w3.org/TR/1999/REC-rdf-syntax-19990222/}  

\bibitem[Stanczyk 1987]{stanczykProcessModellingInformation1987}
Stefan K. Stanczyk (1987): \\
Process modelling for information system
description. \emph{The Open University}, (1987).
\url{https://doi.org/10.21954/ou.ro.0000f821}

\bibitem[Berners-Lee 2005]{rfc3986}
Tim Berners-Lee, Roy T. Fielding, Larry M. Masinter (2005): \\
\textbf{Uniform {Resource Identifier} ({URI}): {Generic Syntax}}.\\
\emph{RFC Editor}
\url{https://doi.org/10.17487/rfc3986}

\bibitem[InfoURIRegistry]{InfoURIRegistry}
(n.d.): "Info" {URI Registry} ({Frozen}). (n.d.).
\url{https://oclc-research.github.io/infoURI-Frozen/} (accessed 24
January 2023)

\bibitem[Juty 2011]{jutyIdentifiersOrgMIRIAM2011}
N. Juty, N. Le Novere, C. Laibe (2011): \\
Identifiers.org and {MIRIAM
Registry}: Community resources to provide persistent identification.
\emph{Nucleic Acids Research}, \textbf{40} (2011) D580--D586.
\url{https://doi.org/10.1093/nar/gkr1097}

\bibitem[Dürst 2005]{rfc3987}
Martin J. Dürst, Michel Suignard (2005): \\
\emph{Internationalized
{Resource Identifiers} ({IRIs})} ({RFC Editor}, 2005).
\url{https://doi.org/10.17487/rfc3987}

\bibitem[Sauermann 2011]{sauermannCoolURIsSemantic2011}
Leo Sauermann, Richard Cyganiak, Max Völkel (2011): \\
\emph{Cool {URIs}
for the semantic web} ({Universität des Saarlandes}, 2011)

\bibitem[w3-owl2-overview]{w3-owl2-overview}
W3C OWL Working Group, ed. (n.d.): {OWL} 2 {Web Ontology Language
Document Overview} ({Second Edition}). (n.d.).
\url{http://www.w3.org/TR/owl2-overview/} (accessed 26 May 2022)

\bibitem[horrocksSemanticWebISWC2002]{horrocksSemanticWebISWC2002}
Ian Horrocks, James Hendler, eds. (2002): \\
\emph{The {Semantic Web} ---
{ISWC} 2002} ({Springer Berlin Heidelberg}, 2002).
\url{https://doi.org/10.1007/3-540-48005-6}

\bibitem[w3-sparql11-overview]{w3-sparql11-overview}
(n.d.): {SPARQL} 1.1 {Overview}. (n.d.).
\url{http://www.w3.org/TR/sparql11-overview/} (accessed 26 May 2022)

\bibitem[w3-rdf11-concepts]{w3-rdf11-concepts}
David Wood, Richard Cyganiak, Markus Lanthaler (2014): \\
{RDF 1.1
Concepts and Abstract Syntax}. (2014).
\url{https://www.w3.org/TR/2014/REC-rdf11-concepts-20140225/}

\bibitem[Goble 2008]{gobleStateNationData2008c}
Carole Goble, Robert Stevens (2008): \\
\textbf{State of the nation in data integration for bioinformatics}.
\emph{Journal of Biomedical Informatics} \textbf{41}
\url{https://doi.org/10.1016/j.jbi.2008.01.008}

\bibitem[Williams 2012]{williamsOpenPHACTSSemantic2012c}
Antony J. Williams, Lee Harland, Paul Groth, Stephen Pettifer, Christine
Chichester, Egon L. Willighagen, Chris T. Evelo, Niklas Blomberg,
Gerhard Ecker, Carole Goble, Barend Mons (2012): \\
Open {PHACTS}:
Semantic interoperability for drug discovery. \emph{Drug Discovery
Today}, \textbf{17} (2012) 1188--1198.
\url{https://doi.org/10.1016/j.drudis.2012.05.016}

\bibitem[Isaac 2009]{w3-skos-primer}
Antoine Isaac and Ed Summers (2009): \\
\textbf{SKOS Simple Knowledge Organization System Primer}. 
W3C Note 
\url{https://www.w3.org/TR/2009/NOTE-skos-primer-20090818/}

\bibitem[Hu 2011]{huHowMatchableAre2011a}
Wei Hu, Jianfeng Chen, Hang Zhang, Yuzhong Qu (2011): \\
\textbf{How matchable are four thousand ontologies on the semantic web}. \\
In Grigoris
Antoniou, Marko Grobelnik, Elena Simperl, Bijan Parsia, Dimitris
Plexousakis, Pieter De Leenheer, Jeff Pan,eds., \emph{The semantic
web: {Research} and applications}, pp. 290--304 \\
ISBN 978-3-642-21033-4

\bibitem[Groth 2014]{grothAPIcentricLinkedData2014b}
Paul Groth, Antonis Loizou, Alasdair J. G. Gray, Carole Goble, Lee
Harland, Steve Pettifer (2014): \\
\textbf{API-centric Linked Data integration: {The Open PHACTS Discovery Platform} case study}.\\
\emph{Journal of Web Semantics}, \textbf{29} \\
\url{https://doi.org/10.1016/j.websem.2014.03.003}

\bibitem[Kamdar 2017]{kamdarSystematicAnalysisTerm2017a}
Maulik R. Kamdar, Tania Tudorache, Mark A. Musen (2017): \\
A systematic
analysis of term reuse and term overlap across biomedical ontologies.
\emph{Semantic Web}, \textbf{8} (2017) 853--871.
\url{https://doi.org/10.3233/sw-160238}

\bibitem[kleinScholarlyContextNot2014a]{kleinScholarlyContextNot2014a}
Martin Klein, Herbert Van de Sompel, Robert Sanderson, Harihar Shankar,
Lyudmila Balakireva, Ke Zhou, Richard Tobin (2014): \\
Scholarly
{Context Not Found}: {One} in {Five Articles Suffers} from {Reference
Rot}. \emph{PLoS ONE}, \textbf{9} (2014) e115253.
\url{https://doi.org/10.1371/journal.pone.0115253}

\bibitem[Polleres 2020]{polleresMoreDecentralizedVision2020a}
Axel Polleres, Maulik Rajendra Kamdar, Javier David Fernández, Tania
Tudorache, Mark Alan Musen (2020): \\
\textbf{A more decentralized vision for linked data}.\\ 
\emph{Semantic Web}, \textbf{11} \\
\url{https://doi.org/10.3233/SW-190380}

\bibitem[Carriero 1010]{carrieroLandscapeOntologyReuse2020a}
Valentina Anita Carriero, Marilena Daquino, Aldo Gangemi, Andrea
Giovanni Nuzzolese, Silvio Peroni, Valentina Presutti, Francesca
Tomasi (2020): \\
\textbf{The landscape of ontology reuse approaches}. \\
In Giuseppe Cota, Marilena Daquino, Gian Luca Pozzato,eds.,
\emph{Applications and practices in ontology design, extraction, and
reasoning} \\
\url{https://doi.org/10.3233/ssw200033}

\bibitem[Guha 2014]{w3-rdf-schema}
Ramanathan Guha and Dan Brickley (2014): \\
\textbf{RDF Schema 1.1}.\\
\emph{W3C} Recommendation \\
\url{http://www.w3.org/TR/rdf-schema/}

\bibitem[Tirmizi 2011]{tirmiziMappingOBOOWL2011a}
Syed Tirmizi, Stuart Aitken, Dilvan A. Moreira, Chris Mungall, Juan
Sequeda, Nigam H. Shah, Daniel P. Miranker (2011): \\
\textbf{Mapping between the {OBO} and {OWL} ontology languages}. \\
\emph{Journal of Biomedical Semantics} \textbf{2}:S3
\url{https://doi.org/10.1186/2041-1480-2-s1-s3}

\bibitem[Berners-Lee 2006]{LinkedDataDesign}
Tim Berners-Lee (2006): \\
Linked {Data} - {Design Issues}. (2006).
\url{https://www.w3.org/DesignIssues/LinkedData.html} (accessed 26 May
2022)

\bibitem[Open Graph]{OpenGraphProtocol}
The {Open Graph} protocol.  
\url{https://ogp.me/}
(accessed 26 May 2022)

\bibitem[Sporny 2015]{w3-rdfa-primer}
Manu Sporny, Ivan Herman, Ben Adida, Mark Birbeck (2015):\\
\textbf{RDFa} 1.1 {Primer} - {Third Edition}}. \\
\url{https://www.w3.org/TR/2015/NOTE-rdfa-primer-20150317/}

\bibitem[WHATWG 2023]{HTMLStandard}
WHATWG (2023): \\
{Microdata}. \\
in: \emph{{HTML Living Standard}}
\url{https://html.spec.whatwg.org/multipage/microdata.html} (accessed 13
June 2023)

\bibitem[Sporny 2020]{w3-json-ld}
Manu Sporny, Dave Longley, Gregg Kellogg, Markus Lanthaler,
Pierre-Antoine Champin, Niklas Lindström (2020): \\
\textbf{JSON-LD 1.1}.\\
\emph{W3C} Recommendation.\\
\url{https://www.w3.org/TR/2020/REC-json-ld11-20200716/}

\bibitem[W3Techs 2023]{UsageStatisticsJSONLD}
W3Techs 2023: \\
\textbf{Usage Statistics of JSON-LD for Websites}. 2023-05. \\
\emph{W3Techs - World Wide Web Technology Surveys}, Q-Success.
\url{https://w3techs.com/technologies/details/da-jsonld} (accessed 18
May 2023)

\bibitem[Verborgh 2018]{DesigningLinkedData2018}
Ruben Verborgh (2018): \\
\textbf{Designing a {Linked Data} developer experience}.\\
\url{https://ruben.verborgh.org/blog/2018/12/28/designing-a-linked-data-developer-experience/}
(accessed 26 May 2022)

\bibitem[w3-shacl]{w3-shacl}
Dimitris Kontokostas, Holger Knublauch (2017):\\
\textbf{Shapes {Constraint Language} ({SHACL})}.\\
\emph{W3C} Recommendation\\
\url{https://www.w3.org/TR/shacl/} (accessed 26 May 2022)

\bibitem[Baker 2019]{ShapeExpressionsShEx}
Thomas Baker and Eric Prud'hommeaux (2019): \\
\textbf{Shape {Expressions} ({ShEx}) 2.1 {Primer}}. \\
\url{http://shex.io/shex-primer/} (accessed 26 May 2022)

\bibitem[Labra Gayo 2017]{gayoValidatingRDFData2017a}
Jose Emilio Labra Gayo, Eric Prud'hommeaux, Iovka Boneva, Dimitris
Kontokostas (2017): \\
\textbf{Validating {RDF Data}}. \\
\emph{Synthesis Lectures on the Semantic Web: Theory and Technology} \textbf{7} \\
\url{https://doi.org/10.2200/s00786ed1v01y201707wbe016}

\bibitem[Thornton 2019]{thorntonUsingShapeExpressions2019a}
Katherine Thornton, Harold Solbrig, Gregory S. Stupp, Jose Emilio Labra
Gayo, Daniel Mietchen, Eric Prud, Andra Waagmeester (2019):
\textbf{Using shape expressions ({ShEx}) to share {RDF} data models and
to guide curation with rigorous validation}. In Pascal Hitzler, Miriam
Fernández, Krzysztof Janowicz, Amrapali Zaveri, Alasdair J.G. Gray,
Vanessa Lopez, Armin Haller, Karl Hammar,eds., \emph{The semantic
web: 16th international conference, {ESWC} 2019, portorož, slovenia,
june 2--6, 2019, proceedings} ({Cham}: {Springer International
Publishing}, 2019), pp. 606--620 ISBN 978-3-030-21347-3

\bibitem[Klímek 2019]{klimekSurveyToolsLinked2019a}
Jakub Klímek, Petr Škoda, Martin Nečaský (2019): \\
\textbf{Survey of tools for linked data consumption}. \\
\emph{Semantic Web} \textbf{10} \\
\url{https://doi.org/10.3233/SW-180316}

\bibitem[Wolstencroft 2011]{wolstencroftRightFieldEmbeddingOntology2011b}
Katy Wolstencroft, Stuart Owen, Matthew Horridge, Olga Krebs, Wolfgang Mueller, Jacky L. Snoep, Franco du Preez, Carole Goble (2011): \\
\textbf{RightField: Embedding ontology annotation in spreadsheets}. \\
\emph{Bioinformatics}, \textbf{27}(14) (2011)
2021--2022.
\url{https://doi.org/10.1093/bioinformatics/btr312}

\bibitem[Page 2011]{pageRESTLinkedData2011}
Kevin R. Page, David C. De Roure, Kirk Martinez (2011): \\
\textbf{{REST} and {Linked Data}}. \\
\emph{Proceedings of the {Second International Workshop} on {RESTful Design} - {WS-REST} '11} 
\url{https://doi.org/10.1145/1967428.1967435}

\bibitem[Bernstein 2016]{bernsteinNewLookSemantic2016a}
Abraham Bernstein, James Hendler, Natalya Noy (2016): \\
\textbf{A new look at the semantic web}. \\
\emph{Communications of the ACM} \textbf{59}(9) \\
\url{https://doi.org/10.1145/2890489}

\bibitem[Delgado 2016]{delgadoInteroperabilityFrameworkDistributed2016a}
José Carlos Martins Delgado (2016): \\
\textbf{An Interoperability Framework and Distributed Platform for Fast Data Applications}.\\
In: \emph{Data {Science} and {Big Data Computing}} \\
\url{https://doi.org/10.1007/978-3-319-31861-5_1}

\bibitem[Zarras 2004]{zarrasComparisonFrameworkMiddleware2004a}
Apostolos Zarras (2004): \\
\textbf{A {Comparison Framework} for {Middleware Infrastructures}}. \\
\emph{The Journal of Object Technology} \textbf{3}(5) \\
\url{https://doi.org/10.5381/jot.2004.3.5.a2}

\bibitem[Bahim 2020]{bahimFAIRDataMaturity2020a}
Christophe Bahim, Carlos Casorrán-Amilburu, Makx Dekkers, Edit Herczog,
Nicolas Loozen, Konstantinos Repanas, Keith Russell, Shelley Stall
(2020): \\
\textbf{The {FAIR} data maturity model: {An} approach to harmonise
{FAIR} assessments}. \\
\emph{Data Science Journal} \textbf{19}(1).\\
\url{https://doi.org/10.5334/dsj-2020-041}

\bibitem[Stallings 1990]{stallingsHandbookComputercommunicationsStandards1990}
William Stallings (1990): \\
\textbf{Handbook of computer-communications standards: {The} open systems ({OSI}) model and {OSI-related}
standards}, 2nd ed. \\
Sams.\\
[ISBN 978-0-672-22697-7](https://identifiers.org/isbn/9780672226977)

\bibitem[Wolstencroft 2013]{wolstencroftTavernaWorkflowSuite2013d}
Katherine Wolstencroft, Robert Haines, Donal Fellows, Alan Williams,
David Withers, Stuart Owen, Stian Soiland-Reyes, Ian Dunlop, Aleksandra
Nenadic, Paul Fisher, Jiten Bhagat, Khalid Belhajjame, Finn Bacall, Alex
Hardisty, Abraham Nieva de la Hidalga, Maria P. Balcazar Vargas, Shoaib
Sufi, Carole Goble (2013): \\
\textbf{The {Taverna} workflow suite: Designing
and executing workflows of {Web Services} on the desktop, web or in the
cloud}. \\
\emph{Nucleic Acids Research} \textbf{41}(W1) W557--W561.\\
\url{https://doi.org/10.1093/nar/gkt328}

\bibitem[Lamprecht 2021]{lamprechtPerspectivesAutomatedComposition2021b}
Anna-Lena Lamprecht, Magnus Palmblad, Jon Ison, Veit Schwämmle, Mohammad
Sadnan Al Manir, Ilkay Altintas, Christopher J. O. Baker, Ammar Ben Hadj
Amor, Salvador Capella-Gutierrez, Paulos Charonyktakis, Michael R.
Crusoe, Yolanda Gil, Carole Goble, Timothy J. Griffin, Paul Groth, Hans
Ienasescu, Pratik Jagtap, Matúš Kalaš, Vedran Kasalica, Alireza
Khanteymoori, Tobias Kuhn, Hailiang Mei, Hervé Ménager, Steffen Möller,
Robin A. Richardson, Vincent Robert, Stian Soiland-Reyes, Robert
Stevens, Szoke Szaniszlo, Suzan Verberne, Aswin Verhoeven, Katherine
Wolstencroft (2021): \\
\textbf{Perspectives on automated composition of workflows
in the life sciences}. \\
\emph{F1000Research} \textbf{10} (2021):897 \\
\url{https://doi.org/10.12688/f1000research.54159.1}

\bibitem[OpenStand 2017]{ModernStandardsParadigm}
OpenStand (2017): \\
The {Modern Standards Paradigm} - {Five Key Principles}.\\
\url{https://open-stand.org/about-us/principles/} (accessed 24 January
2023)

\bibitem[Loo 2022]{looFirstInternationalConference2022}
Tina Loo, ed. (2022): \\
\textbf{First International Conference on FAIR Digital Objects}.\\
\emph{Research Ideas and Outcomes}\\
\url{https://doi.org/10.3897/rio.coll.190}

\bibitem[CNRI 2023]{DOIPAPIHTTPa}
CNRI (2023): \\
\textbf{DOIP API for HTTP Clients} \\
\emph{Cordra® Software Technical Manual Version 2.5.0}\\
Corporation for National Research Initiatives.
\url{https://www.cordra.org/documentation/api/doip-api-for-http-clients.html}
(accessed 13 June 2023)

\bibitem[DOIResolutionDocumentation]{DOIResolutionDocumentation}
(2020): \\
{DOI Resolution Documentation}. (2020).
\url{https://www.doi.org/factsheets/DOIProxy.html} (accessed 24 January
2023)

\bibitem[DOIHandbookResolution]{DOIHandbookResolution}
(2017): \\
\textbf{{DOI Handbook} - {Resolution}}. \emph{{DOI Handbook}}
({International DOI Foundation}, 2017).
\url{https://doi.org/10.1000/182}

\bibitem[Fielding 2000]{fieldingArchitecturalStylesDesign2000a}
Roy Thomas Fielding (2000): \\
\textbf{Architectural styles and the design
of network-based software architectures}\\
Doctoral Thesis, 
\emph{University of California}, Irvine.
\url{https://www.ics.uci.edu/~fielding/pubs/dissertation/top.htm}
(accessed 28 June 2022)

\bibitem[Fielding 2017]{fieldingReflectionsRESTArchitectural2017a}
Roy T. Fielding, Richard N. Taylor, Justin R. Erenkrantz, Michael M.
Gorlick, Jim Whitehead, Rohit Khare, Peyman Oreizy (2017): \\
\textbf{Reflections on the {REST} architectural style and "principled
design of the modern web architecture" (impact paper award)}. \\
\emph{Proceedings of the 2017 11th joint meeting on foundations of software engineering - {ESEC}/{FSE} 2017}, New York, New York, USA.\\
\url{https://doi.org/10.1145/3106237.3121282}

\bibitem[Wittenburg 2023]{cwfr}
Peter Wittenburg, et al (2023): \\
\textbf{Canonical workflow frameworks for research}. \\
\emph{OSF}\\
\url{https://osf.io/3rekv/}

\bibitem[Miller 2021]{OpenAPISpecificationV3}
Darrel Miller, Jeremy Whitlock, Marsh Gardiner, Mike Ralphson, Ron
Ratovsky, Uri Sarid, eds. (2021): \\
\textbf{OpenAPI Specification} v3.1.0.\\
\emph{OpenAPI Initiative}, The Linux Foundation.
\url{https://spec.openapis.org/oas/v3.1.0.html} (accessed 21
March 2023)

\bibitem[schema actions]{SchemaOrgActions}
\textbf{Schema.org {Actions}}.\\
\emph{schema.org}.
\url{https://schema.org/docs/actions.html} (accessed 26 May 2022)

\bibitem[Liu 2007]{w3-wsdl20-primer}
Kevin Liu, David Booth (2007): \\
\textbbf{Web {Services Description Language} ({WSDL})} {Version} 2.0
{Part} 0: {Primer}. \\
\emph{W3C} Recommendation \\
\url{https://www.w3.org/TR/2007/REC-wsdl20-primer-20070626/}

\bibitem[rfc9110]{rfc9110}
Roy T. Fielding, Mark Nottingham, Julian Reschke (2022): \\
\emph{{HTTP
Semantics}} ({RFC Editor}, 2022).
\url{https://doi.org/10.17487/rfc9110}

\bibitem[Speicher 2015]{w3-ldp}
Steve Speicher, John Arwe, Ashok Malhotra, eds. (2015): \\
Linked data
platform 1.0. (2015). \url{https://www.w3.org/TR/2015/REC-ldp-20150226/}

\bibitem[Bonino 2022]{FDOFramework}
Luiz Olavo Bonino da Silva Santos (2022): \\
{FAIR Digital Object Framework
Documentation}. (2022). \url{https://fairdigitalobjectframework.org/}
(accessed 26 May 2022)

\bibitem[w3-prov-o]{w3-prov-o}
(n.d.): {PROV-O}: {The PROV Ontology}. (n.d.).
\url{https://www.w3.org/TR/2013/REC-prov-o-20130430/} (accessed 24
January 2023)

\bibitem[rfc3650]{rfc3650}
Larry Lannom, Lt. Col. Brian P. Boesch, Sam Sun (2003): \\
\emph{Handle
{System Overview}} ({RFC Editor}, 2003).
\url{https://doi.org/10.17487/rfc3650}

\bibitem[Ekuan]{martinekuanWebAPIDesign}
Martin Ekuan, Mick Alberts, Tim Sherer, Udi Dahan, Mike Kistler et
al. (n.d.): Web {API} design best practices - {Azure Architecture
Center}. (n.d.).
\url{https://learn.microsoft.com/en-us/azure/architecture/best-practices/api-design}
(accessed 24 January 2023)

\bibitem[Smith 2007]{smithOBOFoundryCoordinated2007a}
Barry Smith, Michael Ashburner, Cornelius Rosse, Jonathan Bard, William
Bug, Werner Ceusters, Louis J. Goldberg, Karen Eilbeck, Amelia Ireland,
Christopher J. Mungall, Neocles Leontis, Philippe Rocca-Serra, Alan
Ruttenberg, Susanna-Assunta Sansone, Richard H. Scheuermann, Nigam Shah,
Patricia L. Whetzel, Suzanna Lewis (2007): \\
\textbf{The {OBO Foundry}:
Coordinated evolution of ontologies to support biomedical data
integration}. \\
\emph{Nature Biotechnology}, \textbf{25}(11) \\
\url{https://doi.org/10.1038/nbt1346}

\bibitem[Semmler 2022]{DataInformationView}
Tido Semmler, Sergey Danilov, Thomas Rackow, Dmitry Sidorenko, Dirk
Barbi, Jan Hegewald, Dmitri Sein, Qiang Wang, Thomas Jung (2022):\\
\textbf{IPCC DDC: AWI AWI-CM1.1MR model output prepared for CMIP6 CMIP
historical}. \\
\url{https://www.wdc-climate.de/ui/entry?acronym=C6CMAWAWMhi}\\
\url{https://hdl.handle.net/21.14100/2fcf49d3-0608-3373-a47f-0e721b7eaa87}

\bibitem[Albertoni 2023]{w3-vocab-dcat-3}
Riccardo Albertonim David Browning, Simon Cox, Alejandra Gonzalez Beltran, Andrea Perego, Peter Winstanley (2023): \\
\textbf{Data Catalog Vocabulary (DCAT)- Version 3}.\\
\url{https://www.w3.org/TR/2023/WD-vocab-dcat-3-20230307/}


\bibitem[ORESpecificationAbstract]{ORESpecificationAbstract}
Carl Lagoze, 
Herbert Van de Sompel, 
Pete Johnston, 
Michael Nelson, 
Robert Sanderson, 
Simeon Warner (2008):\\
\textbf{ORE Specification} - {Abstract Data Model}. \\
\url{http://www.openarchives.org/ore/1.0/datamodel#Proxies} 

\bibitem[fenselSemanticWebServices2011]{fenselSemanticWebServices2011}
Dieter Fensel, Federico Michele Facca, Elena Simperl, Ioan Toma
(2011): \\
\emph{Semantic {Web Services}} ({Springer Berlin Heidelberg},
2011).
\url{https://doi.org/10.1007/978-3-642-19193-0}

\bibitem[Hausenblas 2012]{OpenData}
Michael Hausenblas et al. (2012): \\
5-star {Open Data}. (2012).
\url{http://5stardata.info/} (accessed 24 January 2023)

\bibitem[Sandvine 2022]{sandvineGlobalInternetPhenomena}
Sandvine (2022): Global {Internet Phenomena Report} 2022.
\url{https://www.sandvine.com/global-internet-phenomena-report-2022}
(accessed 26 May 2022)

\bibitem[Belshe 2022]{rfc7540}
Mike Belshe, Roberto Peon, Martin Thomson (2015): \\
\textbf{Hypertext Transfer Protocol Version 2 (HTTP/2)} \\
\emph{RFC Editor}. RFC 7540
\url{https://doi.org/10.17487/rfc7540}

\bibitem[Bishop 2022]{rfc9114}
Mike Bishop (2022): \\
\textbf{{HTTP}/3}\\
\emph{RFC Editor}. RFC 9114
\url{https://doi.org/10.17487/rfc9114}

\bibitem[Iyengar 2021]{rfc9000}
Jana Iyengar, Martin Thomson (2021): \\
\textbf{{QUIC}: {A UDP-Based Multiplexed} and {Secure Transport}}\\
\emph{RFC Editor}. RFC 9000
\url{https://doi.org/10.17487/rfc9000}

\bibitem[joras2020]{joras2020}
Matt Joras, Yang Chi (2020): \\
\textemph{How {Facebook} is bringing {QUIC} to billions. (2020).
\url{https://engineering.fb.com/2020/10/21/networking-traffic/how-facebook-is-bringing-quic-to-billions}

\bibitem[Fielding 2012]{rfc6570}
Roy T. Fielding, Mark Nottingham, David Orchard, Joe Gregorio, Marc
Hadley (2012): \\
\emph{{URI Template}} ({RFC Editor}, 2012).
\url{https://doi.org/10.17487/rfc6570}

\bibitem[MDN 2023]{ContentNegotiationHTTP}
MDN (2023): \\
\textbf{HTTP Content negotiation}.\\
\emph{Web technology for developers}. \\
MDN Web Docs \\
\url{https://developer.mozilla.org/en-US/docs/Web/HTTP/Content_negotiation}
(accessed 26 May 2022)

\bibitem[Berners-Lee 1998]{berners-lee-cool-uris}
Tim Berners-Lee (1998): \\
Cool {URIs} don't change. (1998).
\url{https://www.w3.org/Provider/Style/URI} (accessed 2 February 2023)

\bibitem[Van de Sompel 2013]{rfc7089}
Herbert Van de Sompel, Michael Nelson, Robert Sanderson (2013):\\
\textbf{{HTTP Framework} for {Time-Based Access} to {Resource States} --
{Memento}}\\
\emph{RFC Editor}, RFC 7089.
\url{https://doi.org/10.17487/rfc7089}

\bibitem[Ellison 2002]{rfc3253}
Tim Ellison, Christopher Kaler, Jim Amsden, Jim Whitehead, Geoffrey
M. Clemm (2002): \\
\textbf{Versioning Extensions to WebDAV} (Web Distributed Authoring and Versioning).\\
\emph{RFC Editor}, RFC 3253.
\url{https://doi.org/10.17487/rfc3253}

\bibitem[Dusseault 2007]{rfc4918}
Lisa M. Dusseault (2007): \\
\textbf{HTTP Extensions for Web Distributed Authoring and Versioning} (WebDAV). \\
\emph{RFC Editor}, RFC 4918.
\url{https://doi.org/10.17487/rfc4918}

\bibitem[Khare 2000]{rfc2817}
Rohit Khare, Scott Lawrence (2000): \\
\emph{Upgrading to {TLS Within HTTP}/1.1}
\emph{RFC Editor}, RFC 2817.
\url{https://doi.org/10.17487/rfc2817}

\bibitem[Sporny 2023]{Draftietfmediamansuffixes00MediaTypes}
Manu Sporny, Amy Guy (2023): \\
\textbf{{Media Types with Multiple Suffixes}} 
Internet Engineering Task Force.
\url{https://datatracker.ietf.org/doc/draft-ietf-mediaman-suffixes/03/}

\bibitem[w3-xmlschema11]{w3-xmlschema11}
Henry Thompson and Sandy Gao and David Beech and Murray Maloney and Noah Mendelsohn and Michael Sperberg-McQueen (2012):\\
\textbf{W3C XML Schema Definition Language} ({XSD}) 1.1 {Part} 1: {Structures}. \\
\emph{W3C} Recommendation\\
\url{https://www.w3.org/TR/2012/REC-xmlschema11-1-20120405/}

\bibitem[Rice 2022]{WebSocketsStandard}
Adam Rice, Ian Hickson, Anne van Kesteren, Yutaka Hirano (2022):
{WebSockets Standard}. (2022). \url{https://websockets.spec.whatwg.org/}
(accessed 26 May 2022)

\bibitem[Capadisli 2017]{w3-ldn}
Sarven Capadisli, Amy Guy, eds. (2017): \\
\textbf{Linked Data Notifications}.\\
\url{https://www.w3.org/TR/2017/REC-ldn-20170502/}

\bibitem[de hÓra 2007]{rfc5023}
Bill de hÓra, Joe Gregorio (2007): \\
\textbf{The {Atom Publishing Protocol}} 
\emph{RFC Editor}, RFC 5023.
\url{https://doi.org/10.17487/rfc5023}

\bibitem[SWORDSpecification]{SWORDSpecification}
Richard Jones and Neil Jefferies (2022): \\
\textbf{SWORD 3.0 Specification}. \\
\url{https://swordapp.github.io/swordv3/swordv3.html} (accessed 26 May
2022)

\bibitem[w3-micropub]{w3-micropub}
Aaron Parecki, ed. (2017): \\
\textbf{Micropub}. (2017).
\url{https://www.w3.org/TR/2017/REC-micropub-20170523/}

\bibitem[Fielding 2014]{rfc7231}
Roy T. Fielding, Julian Reschke (2014): \\
\emph{Hypertext {Transfer Protocol} ({HTTP}/1.1): {Semantics} and {Content}} ({RFC Editor}, 2014).
\url{https://doi.org/10.17487/rfc7231}

\bibitem[HydraW3CCommunity]{HydraW3CCommunity}
(n.d.): Hydra {W3C Community Group}. (n.d.).
\url{https://www.hydra-cg.com/} (accessed 24 January 2023)

\bibitem[Draftkellyjsonhal08]{Draftkellyjsonhal08}
Mike Kelly (2016): \\
\emph{{JSON Hypertext Application Language}}
(Internet Engineering Task Force; Internet Engineering Task Force,
2016). \url{https://datatracker.ietf.org/doc/draft-kelly-json-hal/08/}

\bibitem[RFC8288]{rfc8288}
Mark Nottingham (2017): \\
\emph{Web {Linking}} ({RFC Editor}, 2017).
\url{https://doi.org/10.17487/rfc8288}

\bibitem[RFC7230]{rfc7230}
Roy T. Fielding, Julian Reschke (2014): \\
\emph{Hypertext {Transfer
Protocol} ({HTTP}/1.1): {Message Syntax} and {Routing}} ({RFC Editor},
2014).
\url{https://doi.org/10.17487/rfc7230}

\bibitem[RFC8216]{rfc8216}
Roger Pantos, William May (2017): \\
\emph{{HTTP Live Streaming}} ({RFC
Editor}, 2017).
\url{https://doi.org/10.17487/rfc8216}

\bibitem[iso23009]{iso23009}
(2022): \\
\emph{{ISO}/{IEC} 23009-1:2022 --- information technology ---
dynamic adaptive streaming over HTTP (DASH) --- part 1: Media
presentation description and segment formats} ({ISO}, 2022).
\url{https://www.iso.org/standard/83314.html}

\bibitem[groupFAIRDataMaturity2020]{groupFAIRDataMaturity2020}
FAIR Data Maturity Model Working Group (2020): \\
{FAIR} data maturity
model: {Specification} and guidelines. (2020).
\url{https://doi.org/10.15497/rda00050}

\bibitem[NCBOBioPortal]{NCBOBioPortal}
(n.d.): {NCBO BioPortal}. (n.d.).
\url{https://bioportal.bioontology.org/ontologies} (accessed 26 May
2022)

\bibitem[allcockGlobusStripedGridFTP]{allcockGlobusStripedGridFTP}
W. Allcock, J. Bresnahan, R. Kettimuthu, M. Link (2005): \\
\textbf{The
{Globus Striped GridFTP Framework} and {Server}}. \emph{{SC '05:
Proceedings of the 2005 ACM/IEEE Conference on Supercomputing}}
({Seattle, WA, USA}: {IEEE}, 2005).
\url{https://doi.org/10.1109/sc.2005.72}

\bibitem[ARKIdentifierScheme]{ARKIdentifierScheme}
John A. Kunze, Emmanuelle Bermès (2022): \\
\emph{{The ARK Identifier
Scheme}} (Internet Engineering Task Force; Internet Engineering Task
Force, 2022). \url{https://datatracker.ietf.org/doc/draft-kunze-ark/36/}

\bibitem[rfc3652]{rfc3652}
Larry Lannom, Jason Petrone, Sean Reilly, Sam Sun (2003):
\emph{Handle {System Protocol} (ver 2.1) {Specification}} ({RFC Editor},
2003).
\url{https://doi.org/10.17487/rfc3652}

\bibitem[HandleNetRegistry]{HandleNetRegistry}
(n.d.): Handle.{Net Registry}. (n.d.).
\url{https://www.handle.net/download_hnr.html} (accessed 24 January
2023)

\bibitem[LicensesStandardsOpen]{LicensesStandardsOpen}
(2022): \\
Licenses \& {Standards}. (2022).
\url{https://opensource.org/licenses} (accessed 24 January 2023)

\bibitem[iso16684]{iso16684}
(2019): \\
\emph{{ISO} 16684-1:2019 --- graphic technology --- extensible
metadata platform (XMP) --- part 1: Data model, serialization and core
properties} ({ISO}, 2019). \url{https://www.iso.org/standard/75163.html}

\bibitem[DataW3C]{DataW3C}
(n.d.): Data - {W3C}. (n.d.).
\url{https://www.w3.org/standards/semanticweb/data} (accessed 26 May
2022)

\bibitem[w3-prov-links]{w3-prov-links}
(n.d.): Linking {Across Provenance Bundles}. (n.d.).
\url{https://www.w3.org/TR/2013/NOTE-prov-links-20130430/} (accessed 24
January 2023)

\bibitem[hollandIntroducingRole2014]{hollandIntroducingRole2014}
Vicki Tardif Holland, Jason Johnson (2014): \\
Introducing '{Role}'.
(2014). \url{http://blog.schema.org/2014/06/introducing-role.html}
(accessed 24 January 2023)

\bibitem[SchemaOrg]{SchemaOrg}
(n.d.): Schema.org - {Schema}.org. (n.d.). \url{https://schema.org/}
(accessed 26 May 2022)

\bibitem[DCMIMetadataTerms]{DCMIMetadataTerms}
DCMI Usage Board (2020): \\
{DCMI Metadata Terms}. (2020).
\url{https://www.dublincore.org/specifications/dublin-core/dcmi-terms/2020-01-20/}
(accessed 26 May 2022)

\bibitem[w3-vocab-dcat-2]{w3-vocab-dcat-2}
David Browning, Peter Winstanley, Andrea Perego, Simon Cox, Riccardo
Albertoni, Alejandra Gonzalez Beltran (2020): \\
Data catalog vocabulary
({DCAT}) - version 2. (2020).
\url{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}

\bibitem[FOAFVocabularySpecification]{FOAFVocabularySpecification}
Dan Brickley, Libby Miller (2014): \\
{FOAF Vocabulary Specification}.
(2014). \url{http://xmlns.com/foaf/spec/} (accessed 26 May 2022)

\bibitem[ciccaresePAVOntologyProvenance2013e]{ciccaresePAVOntologyProvenance2013e}
Paolo Ciccarese, Stian Soiland-Reyes, Khalid Belhajjame, Alasdair JG
Gray, Carole Goble, Tim Clark (2013): \\
{PAV} ontology: Provenance,
authoring and versioning. \emph{Journal of Biomedical Semantics},
\textbf{4} (2013) 37.
\url{https://doi.org/10.1186/2041-1480-4-37}

\bibitem[hardistySpecimenDataRefinery2022a]{hardistySpecimenDataRefinery2022a}
Alex Hardisty, Paul Brack, Carole Goble, Laurence Livermore, Ben Scott,
Quentin Groom, Stuart Owen, Stian Soiland-Reyes (2022): \\
The {Specimen
Data Refinery}: {A Canonical Workflow Framework} and {FAIR} {Digital
Object Approach} to {Speeding} up {Digital Mobilisation} of {Natural
History} {Collections}. \emph{Data Intelligence}, \textbf{4} (2022)
320--341.
\url{https://doi.org/10.1162/dint_a_00134}

\bibitem[Bioschemas]{Bioschemas}
Alasdair Gray, Carole Goble, Rafael Jimenez, Bioschemas Community
(2017): \\
\textbf{Bioschemas: From potato salad to protein annotation}. In
Nadeschda Nikitina, Dezhao Song, Achille Fokoue, Peter Haase,eds.,
\emph{Proceedings of the ISWC 2017 posters \& demonstrations and
industry tracks co-located with 16th international semantic web
conference (ISWC 2017)} (Vienna, Austria, 2017).
\url{https://ceur-ws.org/Vol-1963/paper579.pdf}

\bibitem[wieczorekDarwinCoreEvolving2012]{wieczorekDarwinCoreEvolving2012}
John Wieczorek, David Bloom, Robert Guralnick, Stan Blum, Markus Döring,
Renato Giovanni, Tim Robertson, David Vieglais (2012): \\
Darwin {Core}:
{An Evolving Community-Developed Biodiversity Data Standard}. \emph{PLoS
ONE}, \textbf{7} (2012) e29715.
\url{https://doi.org/10.1371/journal.pone.0029715}

\bibitem[EOSC 2016]{10.2777/940154}
European Commission, Directorate-General for Research and Innovation (2016): \\ 
\textbf{Realising the European open science cloud – First report and recommendations of the Commission high level expert group on the European Open Science Cloud}. \\
\emph{Publications Office of the EU} \\
\url{https://doi.org/10.2777/940154}

\bibitem[Norris 2021]{10.1186/s13326-021-00240-6}
Emma Norris, Janna Hastings, Marta M. Marques, Ailbhe N. Finnerty Mutlu,
Silje Zink, Susan Michie (2021): \\
\textbf{Why and how to engage expert
stakeholders in ontology development: Insights from social and
behavioural sciences}. \\
\emph{Journal of Biomedical Semantics} \textbf{12} \\
\url{https://doi.org/10.1186/s13326-021-00240-6}

\bibitem[Jaradeh 201]{10.1007/978-3-030-30760-8_31}
Mohamad Yaser Jaradeh, Allard Oelen, Manuel Prinz, Markus Stocker, Sören Auer (2019): \\
\textbf{Open research knowledge graph: A system walkthrough}. \\
\emph{Digital libraries for open knowledge} 348--351.
\url{https://doi.org/10.1007/978-3-030-30760-8_31}

\bibitem[Fenner 2019]{10.5438/jwvf-8a66}
Martin Fenner, Amir Aryani (2019): \\
\textbf{Introducing the PID graph}. \\
\emph{DataCite Blog}\\
\url{https://doi.org/10.5438/jwvf-8a66}

\bibitem[Singhal 2012]{singhal2012}
Amit Singhal (2012): \\
\textbf{Introducing the knowledge graph: Things, not strings}.
\url{https://blog.google/products/search/introducing-knowledge-graph-things-not/}
(accessed 18 May 2023)

\bibitem[Corcho 2023]{10.48550/arXiv.2305.06746}
Oscar Corcho, Fajar J. Ekaputra, Ivan Heibi, Clement Jonquet, Andras
Micsik, Silvio Peroni, Emanuele Storti (2023): \\
\textbf{A maturity model for catalogues of semantic artefacts}. 
\emph{arXiv} 2305.06746 [cs.DL] \\
\url{https://doi.org/10.48550/arXiv.2305.06746}

\bibitem[Rocca-Serra 2023]{faircookbook}
Philippe Rocca-Serra, Wei Gu, Vassilios Ioannidis, Tooba Abbassi-Daloii,
Salvador Capella-Gutierrez, Ishwar Chandramouliswaran, Andrea
Splendiani, Tony Burdett, Robert T. Giessmann, David Henderson,
Dominique Batista, Ibrahim Emam, Yojana Gadiya, Lucas Giovanni, Egon
Willighagen, Chris Evelo, Alasdair J. G. Gray, Philip Gribbon, Nick
Juty, Danielle Welter, Karsten Quast, Paul Peeters, Tom Plasterer, Colin
Wood, Eelke van der Horst, Dorothy Reilly, Herman van Vlijmen, Serena
Scollen, Allyson Lister, Milo Thurston, Ramon Granell, Gabriel
Backianathan, Sebastian Baier, Anne Cambon Thomsen, Martin Cook, Melanie
Courtot, Mike d'Arcy, Kurt Dauth, Eva Marin del Piico, Leyla Garcia,
Ulrich Goldmann, Valentin Grouès, Daniel J. B. Clarke, Erwan Lefloch,
Isuru Liyanage, Petros Papadopoulos, Cyril Pommier, Emiliano Reynares,
Francesco Ronzano, Alejandra Delfin-Rossaro, Venkata Sagatopam, Ashni
Sedani, Vitaly Sedlyarov, Liubov Shilova, Sukhi Singh, Jolanda Strubel,
Kees van Bochove, Zachary Warnes, Peter Woollard, Fuqi Xu, Andrea
Zaliani, Susanna-Assunta Sansone and (2023): \\
The {FAIR} cookbook -
the essential resource for and by {FAIR} doers. \emph{Scientific Data},
\textbf{10} (2023).
\url{https://doi.org/10.1038/s41597-023-02166-3}

\bibitem[Schultes 2020]{FIP}
Erik Schultes, Barbara Magagna, Kristina Maria Hettne, Robert Pergl,
Marek Suchánek, Tobias Kuhn (2020): \\
\textbf{Reusable {FAIR} implementation profiles as accelerators of {FAIR} convergence}.\\
\emph{International Conference on Conceptual Modeling}, ER 2020: Advances in Conceptual Modeling, 2022-11-03/--06, Vienna, Austria. \\
\emph{Lecture notes in Computer Science} \textbf{12584} \\ 
\url{https://doi.org/10.1007/978-3-030-65847-2_13}

\bibitem[Devaraju 2021]{Devaraju_2021}
Anusuriya Devaraju, Mustapha Mokrane, Linas Cepinskas, Robert Huber,
Patricia Herterich, Jerry de Vries, Vesa Akerman, Hervé L'Hours, Joy
Davidson, Michael Diepenbroek (2021): \\
From conceptualization to
implementation: {FAIR} assessment of research data objects. \emph{Data
Science Journal}, \textbf{20} (2021).
\url{https://doi.org/10.5334/dsj-2021-004}

\bibitem[Wilkinson 2022]{10.5281/zenodo.7463421}
Mark D. Wilkinson, Susanna-Assunta Sansone, Grootveld Marjan, Josefine
Nordling, Richard Dennis, David Hecker (2022): \\
\emph{{FAIR Assessment
Tools: Towards an "Apples to Apples" Comparisons}} ({EOSC} {FAIR}
Metrics subgroup; Zenodo, 2022).
\url{https://doi.org/10.5281/zenodo.7463421}

\bibitem[Verburg 2023]{10.5281/zenodo.7848102}
Maaike Verburg, Robert Huber, Clement Jonquet, Daniel Garijo (2023):
\emph{{FAIR-IMPACT project response to "FAIR Assessment Tools: Towards
an "Apples to Apples" Comparisons"}} ({FAIR-IMPACT}; Zenodo, 2023).
\url{https://doi.org/10.5281/zenodo.7848102}

\bibitem[Berners-Lee 2000]{SemanticWebXML2000}
Tim Berners-Lee (2000): \\
\textbf{Semantic Web on XML}. 
\emph{XML 2000}, Washington DC, 2000-12-06. \\
\url{https://www.w3.org/2000/Talks/1206-xml2k-tbl/slide10-0.html}
(accessed 24 January 2023)

\bibitem[Soiland-Reyes 2022zz]{10.3897/rio.8.e94501}
Stian Soiland-Reyes, Leyla Jael Castro, Daniel Garijo, Marc Portier,
Carole Goble, Paul Groth (2022): \\
Updating linked data practices for
{FAIR} digital object principles. \emph{Research Ideas and Outcomes},
\textbf{8} (2022).
\url{https://doi.org/10.3897/rio.8.e94501}

\bibitem[Soiland-Reyes 2022xx]{10.3233/ds-210053}
Stian Soiland-Reyes, Peter Sefton, Mercè Crosas, Leyla Jael Castro,
Frederik Coppens, José M. Fernández, Daniel Garijo, BjÃ¶rn GrÃ¼ning,
Marco La Rosa, Simone Leo, Eoghan Ó. Carragáin, Marc Portier, Ana
Trisovic, RO-Crate Community, Paul Groth, Carole Goble (2022):
Packaging research artefacts with {RO}-crate. \emph{Data Science},
\textbf{5} (2022) 97--138.
\url{https://doi.org/10.3233/ds-210053}

\bibitem[Soiland-Reyes 2022ff]{10.3897/rio.8.e93937}
Stian Soiland-Reyes, Peter Sefton, Leyla Jael Castro, Frederik Coppens,
Daniel Garijo, Simone Leo, Marc Portier, Paul Groth (2022): \\
Creating
lightweight {FAIR} digital objects with {RO}-crate. \emph{Research Ideas
and Outcomes}, \textbf{8} (2022).
\url{https://doi.org/10.3897/rio.8.e93937}



\end{thebibliography}

